---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to update the backed up data.

# Setup

```{r}
library(rlang,
        include.only = c("%|%", "%||%"))
library(magrittr,
        include.only = c("%>%", "%<>%", "%T>%", "%!>%", "%$%"))

path_backup_rdb <- "data-raw/backups/rdb.rds"
```

# Back up RDB data

NOTE: Full backup might be unreliable since CCM Design changed something under the hood of the API (rate limiting?). See below for a workaround.

```{r}
#| eval: TRUE

rdb:::rfrnds_old(is_draft = NULL,
                 incl_archive = TRUE,
                 use_cache = FALSE) |>
  readr::write_rds(file = path_backup_rdb,
                   compress = "xz",
                   version = 3L)
```

# Update RDB backup

NOTE: Alternative to creating a full backup (chunk above).

```{r}
#| eval: FALSE

data_latest <- rdb:::rfrnds_old(date_time_last_edited_min = clock::as_date(pal::path_mod_time(path_backup_rdb)),
                                is_draft = NULL,
                                incl_archive = TRUE,
                                use_cache = FALSE)

readr::read_rds(path_backup_rdb) |>
  dplyr::filter(!(id %in% data_latest$id)) |>
  dplyr::bind_rows(data_latest) |>
  readr::write_rds(file = path_backup_rdb,
                   compress = "xz",
                   version = 3L)
```

# Back up RDB file attachments

NOTE: Make sure to first commit pending changes to `data-raw/backups/rdb.rds` to ensure file attachments from recently added referendums are backed up, too.

```{r}
path_backup_s3 <- "~/Arbeit/ZDA/Lokal/Projekte/RDB/s3_backup"

rdb::rfrnds(is_draft = NULL,
            use_cache = FALSE) %$%
  files |>
  purrr::list_rbind() |>
  dplyr::filter(!is_deleted) %$%
  s3_object_key |>
  purrr::walk(.f = \(x) {
    if (!fs::file_exists(fs::path(path_backup_s3, x))) {
      rdb::download_file_attachment(s3_object_key = x,
                                    path = path_backup_s3)
    }
  },
  .progress = TRUE)
```
