---
editor_options:
  chunk_output_type: console
---

# NOTES

The C2D Database offers a [privately documented](https://github.com/ccmdesign/c2d-app/blob/master/docs/services.md) API available under `services.c2d.ch` with
the following "endpoints":

-   `referendums`: get data on multiple referendums

    It supports the following URL parameters:

    -   `mode`: the operation mode; one of `stream`, `count` or `search`
    -   `term` the search term used for `mode=search`; note that only the `title$en` field will be searched
    -   `format`: the data format to be returned when `mode=stream`; one of `json` or `csv`
    -   `filter`: a valid [JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents) to customize queries in the
        underlying MongoDB collection; needs to be Base64-encoded; there are various [query
        operators](https://docs.mongodb.com/manual/reference/operator/query/) available to build complex queries

    The following values are used to denote empty values in the returned data:

    -   In integer fields:

        -   `-1L` (means "information not available or nonexistent", an *explicit* `NA`)
        -   `-2L` ([means "information not given"](https://github.com/ccmdesign/c2d-app/issues/13#issuecomment-802466615), an *implicit* `NA`)

    -   In character fields:

        -   `""`
        -   `"-1"`
        -   `"unknown"` (only for certain fields)
        -   `"-2"`
        -   `"not provided"` (only for certain fields)

    -   In array fields:

        -   (Named) `list()`

-   `referendums/ID`: get data on an individual referendum

    `ID` stands for the referendum's unique identifier `oid` (which corresponds to the column `id` in the data returned by the `referendums()` function of this
    package).

-   The returned JSON data has between 22 and 26 fields. Fields that are not always present include:

    -   `archive`
    -   `canton`
    -   `municipality`
    -   `number`

-   `referendums/stats`: get the number of referendums

    It accepts the following URL parameter:

    -   `filter`: a valid [JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents) to customize queries in the
        underlying MongoDB collection; needs to be Base64-encoded; there are various [query
        operators](https://docs.mongodb.com/manual/reference/operator/query/) available to build complex queries

-   `s3_objects`: retrieve a file stored on the C2D [Amazon S3 bucket](https://en.wikipedia.org/wiki/Amazon_S3#Design)

    It's not a real REST API endpoint but rather a URL redirection (I guess). It expects an object key to be appended like this:

    ``` {.http}
    GET https://services.c2d.ch/s3_objects/referendum_5f859fafd1291cc3961f1bc2_0002.pdf
    ```

-   `states`: get the first 10 sub-national states that match the query

    It accepts the following URL parameters to customize the query:

    -   `country`: limits the results to a specific country code
    -   `term`: limits the result to states matching this search term

-   `health`: check API status

# INTERNAL

## Avoid `R CMD check` notes about undefined global objects used in magrittr pipes

Cf. <https://github.com/tidyverse/magrittr/issues/29#issuecomment-74313262>

```{r}
utils::globalVariables(names = c(".",
                                 # tidyselect fns
                                 "all_of",
                                 "any_of",
                                 "matches",
                                 "where",
                                 # other
                                 "applicability_constraint",
                                 "archive",
                                 "committee_name",
                                 "content-disposition",
                                 "country_code",
                                 "country_name",
                                 "date_time_created",
                                 "electorate_abroad",
                                 "electorate_total",
                                 "files",
                                 "id",
                                 "id_official",
                                 "id_sudd",
                                 "id_sudd_prefix",
                                 "is_draft",
                                 "items",
                                 "level",
                                 "municipality",
                                 "remarks",
                                 "result",
                                 "rowid",
                                 "subnational_entity_name",
                                 "sudd_prefix",
                                 "tag_tier_1",
                                 "tag_tier_2",
                                 "tag_tier_3",
                                 "tags_tier_1",
                                 "tags_tier_2",
                                 "tags_tier_3",
                                 "title_de",
                                 "title_fr",
                                 "title_en",
                                 "type",
                                 "url_sudd",
                                 "url_swissvotes",
                                 "value_labels",
                                 "variable_name",
                                 "variable_values",
                                 "votes",
                                 "votes_per_subterritory"))
```

## Package load/unload

```{r}
.onLoad <- function(libname, pkgname) {
  pkgpins::clear(pkg = pkgname,
                 max_age = getOption("c2d.max_cache_lifespan",
                                     default = "30 days"))
}

.onUnload <- function(libpath) {
  pkgpins::deregister(pkg = pkg)
}
```

## Create custom CSS file for package vignettes

Assembles a custom CSS file by combining [rmarkdown's default lightweight vignette
stylesheet](https://github.com/rstudio/rmarkdown/blob/master/inst/rmarkdown/templates/html_vignette/resources/vignette.css) with our custom pkgdown CSS.

The R code chunk below is *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) (`purl = FALSE`) and thus has
to be manually executed in order to regenerate the custom CSS file.

```{r, purl = FALSE}
pal::assert_pkg("brio")

fs::path_package("rmarkdown/templates/html_vignette/resources/vignette.css",
                 package = "rmarkdown") %>%
  brio::read_file() %>%
  paste0("/* BEGIN custom CSS */\n",
         brio::read_file("pkgdown/extra.css"),
         "/* END custom CSS */\n") %>%
  brio::write_file(path = "vignettes/custom.css")
```

## Functions

### `assemble_query_filter`

NOTE: `date` is not stored as an actual [MongoDB `Date` type](https://docs.mongodb.com/manual/reference/bson-types/#std-label-document-bson-type-date) in the
database, so we can't filter on it. It has [been requested](https://github.com/ccmdesign/c2d-app/issues/20) to change this upstream.

```{r}
#' Assemble MongoDB query filter document
#'
#' @param country_code The `country_code`(s) to be included. A character vector.
#' @param subnational_entity_name The `subnational_entity_name`(s) to be included. A character vector.
#' @param municipality The `municipality`(s) to be included. A character vector.
#' @param level The `level`(s) to be included. A character vector.
#' @param type The `type`(s) to be included. A character vector.
#' @param date The `date`(s) to be included. A character vector.
#' @param is_draft `TRUE` means to include only referendum entries with _draft_ status, `FALSE` to include only normal entries. Set to `NULL` in order to
#'   include both draft and normal entries.
#' @param date_time_created_min The minimum `date_time_created` to be included. A [date][base::Date], a [datetime][base::DateTimeClasses], or something
#'   coercible to (like `"2006-01-02"` or `"2006-01-02T15:04:05"`).
#' @param date_time_created_max The maximum `date_time_created` to be included. A [date][base::Date], a [datetime][base::DateTimeClasses], or something
#'   coercible to (like `"2006-01-02"` or `"2006-01-02T15:04:05"`).
#' @param query_filter A valid [MongoDB JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents) which allows for
#'   maximum control over what data is included. This takes precedence over all the other parameters, i.e. if `query_filter` is provided, the parameters
#'   ``r formals(assemble_query_filter) %>% names() %>% setdiff(c("query_filter", "base64_encode")) %>% pal::prose_ls(wrap = "`")`` are ignored.
#' @param base64_encode Whether or not to [Base64](https://en.wikipedia.org/wiki/Base64)-encode the resulting query filter document. Note that the
#'   `query_filter` argument provided to other functions of this package must be Base64-encoded.
#'
#' @return A character scalar containing a valid [MongoDB JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents),
#'   [Base64](https://en.wikipedia.org/wiki/Base64)-encoded if `base64_encode = TRUE`.
#' @keywords internal
assemble_query_filter <- function(country_code = NULL,
                                  subnational_entity_name = NULL,
                                  municipality = NULL,
                                  level = NULL,
                                  type = NULL,
                                  date = NULL,
                                  is_draft = NULL,
                                  date_time_created_min = NULL,
                                  date_time_created_max = NULL,
                                  query_filter = NULL,
                                  base64_encode = TRUE) {
  
  # assemble JSON query filter document if `query_filter` is not provided
  if (is.null(checkmate::assert_string(query_filter, null.ok = TRUE))) {
    
    query_filter <-
      list(country_code =
             country_code %>%
             purrr::map_chr(checkmate::assert_choice,
                            choices = countrycode::codelist$iso2c %>% setdiff(NA_character_),
                            null.ok = TRUE,
                            .var.name = "country_code") %>%
             query_filter_in(),
           canton = query_filter_in(checkmate::assert_character(subnational_entity_name,
                                                                any.missing = FALSE,
                                                                null.ok = TRUE)),
           municipality = query_filter_in(checkmate::assert_character(municipality,
                                                                      any.missing = FALSE,
                                                                      null.ok = TRUE)),
           level =
             level %>%
             purrr::map_chr(checkmate::assert_choice,
                            choices = v_vals("level"),
                            null.ok = TRUE,
                            .var.name = "level") %>%
             purrr::when(length(.) == 0L ~ .,
                         ~ dplyr::recode(., "subnational" = "sub-national") %>% stringr::str_to_sentence()) %>%
             query_filter_in(),
           institution =
             type %>%
             purrr::map_chr(checkmate::assert_choice,
                            choices = v_vals("type"),
                            null.ok = TRUE,
                            .var.name = "type") %>%
             purrr::when(length(.) == 0L ~ .,
                         ~ dplyr::recode("citizens' assembly" = "citizen assembly") %>% stringr::str_to_sentence()) %>%
             query_filter_in(),
           date = query_filter_in(checkmate::assert_character(as.character(date),
                                                              pattern = "\\d{4}-\\d{2}-\\d{2}",
                                                              any.missing = FALSE,
                                                              null.ok = TRUE,
                                                              .var.name = "date")),
           draft = checkmate::assert_flag(is_draft,
                                          null.ok = TRUE),
           created_on = query_filter_date(min = checkmate::assert_posixct(lubridate::as_datetime(date_time_created_min),
                                                                          any.missing = FALSE,
                                                                          .var.name = "date_time_created_min"),
                                          max = checkmate::assert_posixct(lubridate::as_datetime(date_time_created_max),
                                                                          any.missing = FALSE,
                                                                          .var.name = "date_time_created_max"))) %>%
      # remove `NULL` elements
      purrr::compact() %>%
      # convert to base64-encoded JSON
      jsonlite::toJSON(POSIXt = "ISO8601",
                       auto_unbox = TRUE,
                       digits = NA,
                       pretty = FALSE)
  }
  
  query_filter %>% purrr::when(checkmate::assert_flag(base64_encode) ~ jsonlite::base64_enc(.),
                               ~ .)
}
```

### `assert_cols_valid`

```{r}
assert_cols_valid <- function(data,
                              action = cli::cli_abort) {
  
  non_na_col_names <- c("id",
                        "country_code",
                        "date",
                        "level")
  na_col_names <-
    data %>%
    dplyr::select(any_of(non_na_col_names)) %>%
    dplyr::filter(dplyr::if_any(.fns = is.na)) %>%
    dplyr::select(where(~ any(is.na(.x)))) %>%
    colnames()
  n_na_col_names <- length(na_col_names)
  
  if (n_na_col_names) action("Detected {n_na_col_names} column{?s} in {.arg data} that contain forbidden {.val NA}s: {.var {na_col_names}}")
  
  ## check `id`
  if ("id" %in% colnames(data) && anyDuplicated(data$id)) {
    action("Duplicated {.var id}s detected. IDs must be unique.")
  }

  ## check `country_code`
  if ("country_code" %in% colnames(data)) {
    
    choices <- countrycode::codelist$iso2c %>% setdiff(NA_character_)
    check <- checkmate::check_subset(levels(data$country_code),
                                     choices = choices)
    if (!isTRUE(check)) action("Failed to validate {.var levels(country_code)}. {check}")
    
    ## ensure `position_government` is present when `country_code = "CH" & level = "national"`
    if ((data %>%
         dplyr::filter(country_code == "CH" & level == "national") %>%
         nrow() %>%
         magrittr::is_greater_than(0L))
        && !("position_government" %in% colnames(data))) {
      
      action(paste0("Referendums with {.code country_code = \"CH\" & level = \"national\"} present in {.arg data} but column {.var ",
                    "position_government} is missing."))
    }
  }
  
  ## check `date`
  if ("date" %in% colnames(data)) {
    
    check <- checkmate::check_date(data$date,
                                   any.missing = FALSE)
    if (!isTRUE(check)) action("Failed to validate {.var date}. {check}")
  }
  
  ## check `level`
  if ("level" %in% colnames(data)) {
    
    choices <- v_vals("level")
    check <- checkmate::check_subset(levels(data$level),
                                     choices = choices)
    if (!isTRUE(check)) action("Failed to validate {.var levels(level}. {check}")
  }
  
  ## check `subnational_entity_name`
  ## TODO: check `subnational_entity_code` instead once it's available
  if (any(data[["level"]] != "national")) {
    
    if (!("subnational_entity_name" %in% colnames(data))) {
      action(paste0("Referendums of {.var level} below {.val national} present in {.arg data} but column {.var subnational_entity_name} is missing."))
    }
    
    ix_missing_subnational_entities <-
      data %>%
      tibble::rowid_to_column() %>%
      dplyr::filter(level != "national" & is.na(subnational_entity_name)) %$%
      rowid
    
    n_missing_subnational_entities <- length(ix_missing_subnational_entities)
    
    if (n_missing_subnational_entities) {
      action(paste0("{n_missing_subnational_entities} row{?s} in {.arg data} {?is/are} missing a {.var subnational_entity_name}. Affected {?is/are} ",
                    "the row{?s} with ind{?ex/ices} {.val {ix_missing_subnational_entities}}."))
    }
  }
  
  ## check `municipality`
  if (any(data[["level"]] == "local")) {
    
    if (!("municipality" %in% colnames(data))) {
      action(paste0("Referendums of {.var level = \"local\"} present in {.arg data} but column {.var municipality} is missing."))
    }
    
    ix_missing_municipalities <-
      data %>%
      tibble::rowid_to_column() %>%
      dplyr::filter(level == "local" & is.na(municipality)) %$%
      rowid
    
    n_missing_municipalities <- length(ix_missing_municipalities)
    
    if (n_missing_municipalities) {
      action(paste0("{n_missing_municipalities} row{?s} in {.arg data} {?is/are} missing a {.var municipality}. Affected {?is/are} the row{?s} with ",
                    "ind{?ex/ices} {.val {ix_missing_subnational_entities}}."))
    }
  }
  
  ## check variables that are only meant to be set for Swiss national referendums
  ## TODO: Remove this as soon as [issue #52](https://github.com/ccmdesign/c2d-app/issues/52) is resolved.
  ### `votes_per_subterritory`
  if (all(c("votes_per_subterritory", "level", "country_code") %in% colnames(data))) {
    
    ix_illegal_votes_per_subterritory <-
      data %>%
      tibble::rowid_to_column() %>%
      dplyr::filter((level != "national" | country_code != "CH") & !(votes_per_subterritory %>% purrr::map_lgl(is.null))) %$%
      rowid
    
    n_illegal_votes_per_subterritory <- length(ix_illegal_votes_per_subterritory)
    
    if (n_illegal_votes_per_subterritory) {
      action(paste0("{n_illegal_votes_per_subterritory} row{?s} in {.arg data} {?has/have} a {.var position_government} set although they are not Swiss ",
                    "national referendums. Affected {?is/are} the row{?s} with ind{?ex/ices} {.val {ix_illegal_votes_per_subterritory}}."))
    }
  }
  ### non-list vx
  c("lower_house_yes",
    "lower_house_no",
    "lower_house_abstentions",
    "upper_house_yes",
    "upper_house_no",
    "upper_house_abstentions",
    "position_government") %>%
    purrr::walk(function(v_name) {
      
      if (all(c(v_name, "level", "country_code") %in% colnames(data))) {
        
        ix_illegal <-
          data %>%
          tibble::rowid_to_column() %>%
          dplyr::filter((level != "national" | country_code != "CH") & !is.na(!!as.symbol(v_name))) %$%
          rowid
        
        n_illegal <- length(ix_illegal)
        
        if (n_illegal) {
          action(paste0("{n_illegal} row{?s} in {.arg data} {?has/have} a {.var {v_name}} set although {cli::qty(n_illegal)}{?it is not a/they are not} Swiss",
                        " national referendum{?s}. Affected {?is/are} the row{?s} with ind{?ex/ices} {.val {ix_illegal}}."))
        }
      }
    })
  
  invisible(data)
}
```

### `auth_session`

TODO: Tweak token expiration time [once we know it](https://github.com/ccmdesign/c2d-app/issues/47).

```{r}
#' Authenticate a user session for the [C2D API](https://github.com/ccmdesign/c2d-app/blob/master/docs/services.md#1-reflexive-routes)
#'
#' Creates a new user session token if necessary. The token is stored in the R option `c2d.user_session_tokens`, a [tibble][tibble::tbl_df] with the columns
#' `email`, `token` and `date_time_last_active`.
#' 
#' By default, the `email` and `password` are read from the first element of the R option `c2d.credentials` (a named character vector where names are e-mail
#' addresses and values are passwords).
#'
#' @param email The e-mail address of the user for which a session should be created. A character scalar.
#' @param password The password of the user for which a session should be created. A character scalar.
#'
#' @return The user session token as a character scalar, invisibly.
#' @keywords internal
auth_session <- function(email = names(getOption("c2d.credentials")[1L]),
                         password = getOption("c2d.credentials")[1L]) {
  
  # get existing tokens or initialize empty tibble
  tokens <-
    getOption("c2d.user_session_tokens") %>%
    purrr::when(all(c("email", "token", "date_time_last_active") %in% colnames(.)) ~ .,
                ~ tibble::tibble(email = character(),
                                 token = character(),
                                 date_time_last_active = lubridate::as_datetime(NULL)))
  
  # extract latest token
  token <- tokens %>% dplyr::filter(email == !!email)
  if (nrow(token)) token %<>% dplyr::filter(date_time_last_active == max(date_time_last_active))
  
  # ensure token is not expired (checked if older than 1 day), else set to `NULL`
  if (nrow(token) &&
      checkmate::test_string(token$token, min.chars = 1L) &&
      ((token$date_time_last_active > (lubridate::now() - lubridate::duration(1L, units = "days"))) || !is_session_expired(token$token))) {
    
    token <- token$token
    
  } else {
    
    token <- NULL
  }
  
  # create new session if necessary
  if (is.null(token)) {
    
    pal::cli_process_expr(msg = "Authenticating new user session",
                          expr = {
                            
                            token <-
                              httr::RETRY(verb = "POST",
                                          url = "https://services.c2d.ch/users/session",
                                          times = 5L,
                                          encode = "json",
                                          body = list(email = checkmate::assert_string(email),
                                                      password = checkmate::assert_string(password))) %>%
                              # ensure we actually got a JSON response
                              pal::assert_mime_type(mime_type = "application/json",
                                                    msg_suffix = mime_error_suffix) %>%
                              # parse response
                              httr::content(as = "parsed") %$%
                              token
                          })
  }
  
  # update `c2d.user_session_tokens` option
  options(c2d.user_session_tokens =
            tokens %>%
            dplyr::filter(token != !!token) %>%
            tibble::add_row(email = email,
                            token = token,
                            date_time_last_active = lubridate::now()))
  # return token
  invisible(token)
}
```

### `drop_disabled_vx`

```{r}
drop_disabled_vx <- function(data,
                             to_drop) {
  
  to_drop_present <- intersect(to_drop, colnames(data))
  n_to_drop_present <- length(to_drop_present)
  
  if (n_to_drop_present) {
    
    cli::cli_alert_warning(paste0("The {cli::qty(n_to_drop_present)} column{?s} {.var {to_drop_present}} in {.arg data} are ignored because setting/altering ",
                                  "the corresponding values is disabled."))
    
    data %<>% dplyr::select(-any_of(to_drop))
  }
  
  data
}
```

### `flatten_array_as_is`

A helper function to convert an R list to JSON using `jsonlite::toJSON(auto_unbox = TRUE)`.

```{r}
flatten_array_as_is <- function(x) {
  
  unlist(x) %>% purrr::when(is.null(.) ~ .,
                            ~ I(.))
}
```

### `httr_config`

[Before 2021-05-25](https://github.com/ccmdesign/c2d-app/issues/10), the C2D API server didn't provide the [intermediate *R3*
certificate](https://crt.sh/?q=+730c1bdcd85f57ce5dc0bba733e5f1ba5a925b2a771d640a26f7a454224dad3b) (issued by Let's Encrypt) and since [curl doesn't support
*Authority Information Access* yet](https://github.com/curl/curl/issues/2793) to automatically discover it, we had to manually specify it. Thus, the certificate
was shipped with this package (found under `certs/3479778542.crt`). Because this isn't needed anymore, the cert as well as the code chunk below (`purl = FALSE`)
are ignored from package builds but kept in case they are needed again at some point in the future.

```{r, purl = FALSE}
httr_config <- function() {
  
  httr::config(cainfo = fs::path_package(package = pkg,
                                         "certs", "3479778542.crt"))
}
```

### `is_session_expired`

```{r}
is_session_expired <- function(token) {
  
  response <-
    httr::RETRY(verb = "GET",
              url = "https://services.c2d.ch/users/profile",
              config = httr::add_headers(Authorization = paste("Bearer", token),
                                         Connection = "keep-alive"),
              times = 5L) %>%
    # ensure we actually got a JSON response
    pal::assert_mime_type(mime_type = "application/json",
                          msg_suffix = mime_error_suffix) %>%
    # parse response
    httr::content(as = "parsed")
  
  is.null(response[["profile"]])
}
```

### `lower_non_abbrs`

```{r}
#' Lowercase non-abbreviations
#'
#' @param x A character vector.
#'
#' @return A character vector of the same lenght as `x`.
#' @keywords internal
lower_non_abbrs <- function(x) {
  
  x %>%
    stringr::str_split(pattern = "\\s") %>%
    purrr::map_chr(~ .x %>%
                     dplyr::if_else(stringr::str_detect(string = .,
                                                        pattern = "^[^[:lower:]]+$"),
                                    .,
                                    stringr::str_to_lower(.)) %>%
                     paste0(collapse = " "))
}
```

### `read_toml`

```{r}
#' Read in a TOML file
#'
#' @param path Path to the TOML file. A [path][fs::fs_path] or something coercible to.
#'
#' @return The TOML file's content as a [strict list][xfun::strict_list].
#' @keywords internal
read_toml <- function(path) {
  
  pal::assert_pkg("RcppTOML")
  pal::assert_pkg("xfun")
  
  path %>%
    purrr::when(length(.) > 0L ~ RcppTOML::parseTOML(input = .,
                                                     escape = FALSE),
                ~ NULL) %>%
    xfun::as_strict_list()
}
```

### `rename_from_list`

NOTE that an alternative approach reducing `names_list` to the ones actually present and then `dplyr::rename(!!!names_list_present)` is very inefficient.

```{r}
rename_from_list <- function(x,
                             names_list) {
  if (length(x)) {
    
    names_new <- names(x) %>% dplyr::recode(!!!names_list)
    
    x %>% magrittr::set_names(names_new)
    
  } else {
    x
  }
}
```

### `tidy_date`

```{r}
tidy_date <- function(x) {
  
  x %>%
    unlist(use.names = FALSE) %>%
    magrittr::divide_by(1000L) %>%
    lubridate::as_datetime()
}
```

### `tidy_referendums`

TODO:

-   Upstream renamings and recodings as far as possible and subsequently remove them from `tidy_referendums()`, `assemble_query_filter()` and all other affected
    places.

```{r}
#' Tidy "raw" C2D API referendum data
#'
#' Converts the "raw" MongoDB data from the C2D API to the tidied [referendums()] schema.
#'
#' You can reverse this function again using [untidy_referendums()].
#'
#' @param data The MongoDB data as a list (converted from the JSON returned by the C2D API using [jsonlite::fromJSON()]).
#' @param tidy Whether or not to tidy the referendum data, i.e. apply various data cleansing tasks and add additional variables. If `FALSE`, the raw MongoDB
#'   referendum data will only be modified just enough to be able to return it as a tibble. Note that the latter doesn't conform to the 
#'   [codebook][data_codebook] (i.a. different variable names).
#'
#' @return `r pkgsnip::return_label("data")`
#' @keywords internal
tidy_referendums <- function(data,
                             tidy = TRUE) {
  
  checkmate::assert_flag(tidy)
  
  data %<>%
    # unnest columns and ensure list type for multi-value columns
    # NOTE that `tidyr::unnest()` is unbearably slow, so we use a custom function
    purrr::map(.f = function(l,
                             category_names = names(l$categories),
                             context_names = names(l$context),
                             title_langs = names(l$title)) {
      
      for (name in category_names) {
        l[[paste0("categories.", name)]] <- l$categories[[name]]
      }
      
      for (name in context_names) {
        l[[paste0("context.", name)]] <- l$context[[name]]
      }
      
      for (lang in title_langs) {
        l[[paste0("title.", lang)]] <- l$title[[lang]]
      }
      
      l$categories <- NULL
      l$context <- NULL
      l$title <- NULL
      
      for (name in c("tags",
                     "categories.action",
                     "categories.special_topics",
                     "categories.excluded_topics")) {
        l[[name]] %<>% purrr::flatten_chr() %>% list()
      }
      
      for (name in c("archive",
                     "files",
                     "context.votes_per_canton")) {
        l[[name]] %<>% list()
      }
      
      l
    }) %>%
    # convert to tibble
    purrr::map_dfr(tibble::as_tibble_row)
  
  if (nrow(data)) {
    
    # check data integrity
    ## ensure mandatory fields are never `NA`
    if (data %>%
        dplyr::select(all_of(referendum_fields$required)) %>%
        dplyr::filter(dplyr::if_any(.fns = is.na)) %>%
        nrow()) {
      
      cli::cli_abort("Mandatory fields detected that contain {.val NA} values. Please debug.")
    }
    
    # tidy data
    if (tidy) {
      
      data %<>%
        # rename variables (mind that the MongoDB-based API doesn't demand a schema)
        rename_from_list(names_list = v_names) %>%
        # add vars which aren't always included
        dplyr::full_join(y = tibble::tibble(subnational_entity_name = character(),
                                            municipality = character(),
                                            id_sudd = character()),
                         by = intersect(c("subnational_entity_name",
                                          "municipality",
                                          "id_sudd"),
                                        colnames(.))) %>%
        
        # create/recode variables
        dplyr::mutate(
          # ensure all supposed to floating-point numbers are actually of type double (JSON API is not reliable in this respect)
          dplyr::across(any_of(c("subterritories_no",
                                 "subterritories_yes",
                                 "date_time_created")),
                        as.double),
          
          # use explicit NA values
          dplyr::across(where(is.integer),
                        ~ dplyr::if_else(.x %in% c(-1L, -2L),
                                         NA_integer_,
                                         .x)),
          dplyr::across(where(is.character),
                        ~ dplyr::if_else(.x %in% c("", "-1", "-2"),
                                         NA_character_,
                                         .x)),
          dplyr::across(any_of(c("subterritories_yes", "subterritories_no")),
                        ~ dplyr::if_else(.x %in% c(-1.0, -2.0),
                                         NA_real_,
                                         .x)),
          result = result %>% dplyr::if_else(. %in% c("Unknown", "Not provided"),
                                             NA_character_,
                                             .),
          
          # convert all values to lowercase
          ## vectors
          dplyr::across(c(level,
                          result,
                          type,
                          any_of(c("inst_legal_basis_type",
                                   "inst_object_type",
                                   "inst_object_legal_level",
                                   "inst_object_revision_extent",
                                   "inst_trigger_type",
                                   "inst_trigger_actor_level",
                                   "inst_trigger_time_limit",
                                   "inst_quorum_approval",
                                   "inst_precondition_decision"))),
                        stringr::str_to_lower),
          ## lists
          dplyr::across(.cols = any_of(c("inst_object_revision_modes",
                                         "inst_topics_only",
                                         "inst_topics_excluded")),
                        .fns = purrr::map,
                        .f = stringr::str_to_lower),
          
          # convert non-abbreviations only to lowercase
          dplyr::across(.cols = any_of(c("inst_object_author",
                                         "inst_trigger_actor",
                                         "inst_precondition_actor")),
                        .fns = purrr::map_chr,
                        .f = lower_non_abbrs),
          
          # specific recodings
          ## binary (dummies)
          dplyr::across(any_of("position_government"),
                        ~ dplyr::case_when(.x == "Acceptance" ~ "yes",
                                           .x == "Rejection" ~ "no",
                                           TRUE ~ NA_character_)),
          dplyr::across(any_of("inst_has_urgent_legal_basis"),
                        ~ dplyr::case_when(.x == "Urgent" ~ TRUE,
                                           .x == "Normal" ~ FALSE,
                                           TRUE ~ NA)),
          dplyr::across(any_of("inst_is_binding"),
                        ~ dplyr::case_when(.x == "Binding" ~ TRUE,
                                           .x == "Non-binding" ~ FALSE,
                                           TRUE ~ NA)),
          dplyr::across(any_of("inst_is_counter_proposal"),
                        ~ dplyr::case_when(.x == "Yes" ~ TRUE,
                                           .x == "No" ~ FALSE,
                                           TRUE ~ NA)),
          dplyr::across(any_of("inst_is_assembly"),
                        ~ dplyr::case_when(.x == "Assembly" ~ TRUE,
                                           .x == "Ballot" ~ FALSE,
                                           TRUE ~ NA)),
          dplyr::across(any_of("inst_has_precondition"),
                        ~ dplyr::case_when(.x == "Exists" ~ TRUE,
                                           .x == "Does not exist" ~ FALSE,
                                           TRUE ~ NA)),
          ## nominal
          ### flatten `id`
          id = purrr::flatten_chr(id),
          ### extract `id_official` from `id_sudd` (the latter consists of a two-letter country code plus a 6-digit number)
          id_official =
            id_sudd %>%
            dplyr::if_else(stringr::str_detect(string = .,
                                               pattern = "^\\D"),
                           NA_character_,
                           .),
          id_sudd =
            id_sudd %>%
            dplyr::if_else(stringr::str_detect(string = .,
                                               pattern = "^\\D"),
                           .,
                           NA_character_) %>%
            # everything beyond the 8th char seems to be manually added -> strip!
            stringr::str_sub(end = 8L),
          ### split `tags` into separate per-tier vars
          tags_tier_1 = tags %>% purrr::map(infer_tags,
                                            tier = 1L),
          tags_tier_2 = tags %>% purrr::map(infer_tags,
                                            tier = 2L),
          tags_tier_3 = tags %>% purrr::map(~ .x[.x %in% tags(tiers = 3L)]),
          ### various cleanups
          level = stringr::str_replace(string = level,
                                       pattern = "sub-national",
                                       replacement = "subnational"),
          type = type %>% dplyr::recode("citizen assembly" = "citizens' assembly",
                                        "not provided" = NA_character_),
          dplyr::across(any_of(c("inst_trigger_actor",
                                 "inst_object_author")),
                        dplyr::recode,
                        "institution" = "other institution"),
          dplyr::across(any_of("inst_object_type"),
                        dplyr::recode,
                        "legal text (ausformulierter vorschlag)" = "legal text (formulated proposal)",
                        "legal text (allg. anregung)" = "legal text (general proposal)"),
          dplyr::across(any_of("inst_object_revision_divisibility"),
                        dplyr::recode,
                        "Whole text only" = "as a whole",
                        "Splitting up possible" = "split up",
                        "Variants possible" = "as variants",
                        "Variants / splitting up possible" = "both split up and as variants"),
          dplyr::across(.cols = any_of("inst_topics_only"),
                        .fns = purrr::map,
                        .f = dplyr::recode,
                        "infrastructural act" = "infrastructural acts",
                        "competence shift" = "competence shifts",
                        "financial act" = "financial acts",
                        "financial act (expenses)" = "financial acts (expenses)",
                        "financial act (taxes)" = "financial acts (taxes)",
                        "financial act (obligations)" = "financial acts (obligations)",
                        "total revision of the constitution" = "total revisions of the constitution"),
          dplyr::across(.cols = any_of("inst_topics_excluded"),
                        .fns = purrr::map,
                        .f = dplyr::recode,
                        "budget" = "budgets",
                        "parliamentary competence" = "everything within parliamentary competence"),
          dplyr::across(any_of("inst_quorum_turnout"),
                        stringr::str_replace_all,
                        pattern = c("^(\\s+)?>(\\s+)?" = ">\u202f",
                                    "(\\s+)?%(\\s+)?$" = "\u202f%")),
          ## ordinal
          ## interval
          date = lubridate::as_date(date),
          date_time_created = tidy_date(date_time_created),
          ## undefined
          files = files %>% purrr::map(~ .x %>% purrr::map(~ .x %>%
                                                             # unnest and restore `date`
                                                             purrr::assign_in(where = "date",
                                                                              value = tidy_date(.$date)) %>%
                                                             # change subvariable names
                                                             rename_from_list(names_list = sub_v_names$files))),
          votes_per_subterritory = votes_per_subterritory,
          
          # variable creations
          url_sudd = dplyr::if_else(!is.na(id_sudd),
                                    paste0("https://sudd.ch/event.php?id=", id_sudd),
                                    NA_character_),
          url_swissvotes = dplyr::if_else(country_code == "CH" & level == "national" & !is.na(id_official),
                                          paste0("https://swissvotes.ch/vote/", id_official),
                                          NA_character_)
        ) %>%
        
        # remove obsolete vars
        dplyr::select(-tags) %>%
        
        # convert to (ordered) factor where appropriate
        ## based on codebook
        dplyr::mutate(dplyr::across(.fns = ~ {
          
          metadata <- data_codebook %>% dplyr::filter(variable_name == dplyr::cur_column())
          if (nrow(metadata) != 1L) cli::cli_abort("Missing codebook metadata! Please debug")
          
          if (!is.logical(.x) && !is.null(metadata$variable_values[[1L]])) {
            
            lvls <- purrr::flatten_chr(metadata$variable_values)
            is_ordered <- metadata$value_scale %in% c("ordinal_ascending", "ordinal_descending")
            
            if (is.list(.x)) {
              .x %>% purrr::map(.f = factor,
                                levels = lvls,
                                ordered = is_ordered)
            } else {
              factor(x = .x,
                     levels = lvls,
                     ordered = is_ordered)
            }
          } else .x
        })) %>%
        ## nominal vars without variable_values in codebook
        dplyr::mutate(dplyr::across(any_of(c("country_code",
                                             "country_name",
                                             "subnational_entity_name",
                                             "municipality")),
                                    as.factor)) %>%
        # add variable labels
        labelled::set_variable_labels(.labels = var_lbls,
                                      .strict = FALSE) %>%
        # reorder columns
        dplyr::relocate(id,
                        id_official,
                        id_sudd,
                        country_code,
                        country_name,
                        subnational_entity_name,
                        municipality,
                        level,
                        date,
                        title_de,
                        title_fr,
                        title_en,
                        committee_name,
                        result,
                        any_of("subterritories_yes"),
                        any_of("subterritories_no"),
                        any_of("electorate_total"),
                        any_of("electorate_abroad"),
                        any_of("votes_yes"),
                        any_of("votes_no"),
                        any_of("votes_empty"),
                        any_of("votes_invalid"),
                        any_of("votes_per_subterritory"),
                        any_of("lower_house_yes"),
                        any_of("lower_house_no"),
                        any_of("lower_house_abstentions"),
                        any_of("upper_house_yes"),
                        any_of("upper_house_no"),
                        any_of("upper_house_abstentions"),
                        any_of("position_government"),
                        tags_tier_1,
                        tags_tier_2,
                        tags_tier_3,
                        remarks,
                        files,
                        url_sudd,
                        url_swissvotes,
                        is_draft,
                        date_time_created,
                        archive,
                        type,
                        any_of("inst_legal_basis_type"),
                        any_of("inst_has_urgent_legal_basis"),
                        any_of("inst_is_binding"),
                        any_of("inst_is_counter_proposal"),
                        any_of("inst_is_assembly"),
                        any_of("inst_trigger_type"),
                        any_of("inst_trigger_actor"),
                        any_of("inst_trigger_actor_level"),
                        any_of("inst_trigger_threshold"),
                        any_of("inst_trigger_time_limit"),
                        any_of("inst_object_type"),
                        any_of("inst_object_author"),
                        any_of("inst_object_legal_level"),
                        any_of("inst_object_revision_extent"),
                        any_of("inst_object_revision_modes"),
                        any_of("inst_object_revision_divisibility"),
                        any_of("inst_topics_only"),
                        any_of("inst_topics_excluded"),
                        any_of("inst_quorum_turnout"),
                        any_of("inst_quorum_approval"),
                        any_of("inst_has_precondition"),
                        any_of("inst_precondition_actor"),
                        any_of("inst_precondition_decision"))
      
      # correct known errors
      ## `id_sudd`
      data$id_sudd[data$country_code == "AZ" & data$id_sudd == "az021009"] <- "az022009"
      data$id_sudd[data$country_code == "BS" & data$id_sudd == "bs021002"] <- "bs022002"
      data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011972"] <- "mg011972"
      data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011975"] <- "mg011975"
      data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011992"] <- "mg011992"
      data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011995"] <- "mg011995"
      data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011998"] <- "mg011998"
      data$id_sudd[data$country_code == "MG" & data$id_sudd == "md012007"] <- "mg012007"
      data$id_sudd[data$country_code == "MG" & data$id_sudd == "md012010"] <- "mg012010"
      data$id_sudd[data$country_code == "MH" & data$id_sudd == "mh00578"] <- "mh011979"
      data$id_sudd[data$country_code == "MW" & data$id_sudd == "mw012003"] <- "mw011993"
      data$id_sudd[data$country_code == "EC" & data$id_sudd == "ec031993"] <- "ec031994"
      data$id_sudd[data$country_code == "EC" & data$id_sudd == "ec021006"] <- "ec022006"
      data$id_sudd[data$country_code == "GQ" & data$id_sudd == "gq012968"] <- "gq011968"
      data$id_sudd[data$country_code == "GT" & data$id_sudd == "gt011998"] <- "gt011999"
      data$id_sudd[data$country_code == "GT" & data$id_sudd == "gt021998"] <- "gt021999"
      data$id_sudd[data$country_code == "GT" & data$id_sudd == "gt031998"] <- "gt031999"
      data$id_sudd[data$country_code == "LI" & data$id_sudd == "li021920"] <- "li021930"
      data$id_sudd[data$country_code == "LT" & data$id_sudd == "lt031993"] <- "lt031994"
      data$id_sudd[data$country_code == "NF" & data$id_sudd == "nf011980"] <- "nf011979"
      data$id_sudd[data$country_code == "TV" & data$id_sudd == "tv011994"] <- "tv011974"
      data$id_sudd[data$country_code == "UY" & data$id_sudd == "uy011918"] <- "uy011917"
      data$id_sudd[data$country_code == "YE" & data$id_sudd == "ye012991"] <- "ye011991"
      # data$id_sudd[data$country_code == "" & data$id_sudd == ""] <- ""
    }
  }
  
  # convert nested list cols to tibbles
  data %>% dplyr::mutate(dplyr::across(any_of(c("context.votes_per_canton",
                                                "files",
                                                "votes_per_subterritory")),
                                       ~ .x %>% purrr::map(~ { if (length(.x)) .x %>% purrr::map_dfr(tibble::as_tibble) else NULL })),
                         dplyr::across(any_of("archive"),
                                       ~ .x %>% purrr::map(~ { if (length(.x)) tibble::as_tibble(.x) else NULL })))
}
```

### `untidy_date`

```{r}
untidy_date <- function(x) {
  
  as.numeric(x) %>%
    magrittr::multiply_by(1000) %>%
    as.list() %>%
    magrittr::set_names(rep("$date",
                            times = length(.)))
}
```

### `untidy_referendums`

```{r}
#' Untidy into "raw" C2D API referendum data
#'
#' Converts from the tidied [referendums()] to the "raw" MongoDB schema used by the C2D API. Basically reverts [tidy_referendums()].
#'
#' @param data The data to untidy as returned by [referendums()].
#'
#' @return
#' If `as_tibble = FALSE`, a list with one element per referendum, suitable to convert [jsonlite::toJSON()] and then feed to the C2D API.
#'
#' Otherwise a [tibble][tibble::tbl_df] of the same format as returned by [`referendums(tidy = FALSE)`][referendums].
#' @keywords internal
untidy_referendums <- function(data,
                               as_tibble = FALSE) {
  
  checkmate::assert_flag(as_tibble)
  v_names_inverse <- names(v_names) %>% magrittr::set_names(purrr::flatten_chr(v_names))
  sub_v_names_files_inverse <- names(sub_v_names$files) %>% magrittr::set_names(purrr::flatten_chr(sub_v_names$files))
  
  # restore `number`
  if (all(c("id_official", "id_sudd") %in% colnames(data))) {
    data %<>% dplyr::mutate(id_sudd = dplyr::if_else(is.na(id_sudd),
                                                     id_official,
                                                     id_sudd))
  }
  
  data %<>%
    dplyr::mutate(
      # restore strings
      dplyr::across(c(any_of("date"),
                      where(is.factor)),
                    as.character),
      dplyr::across(where(is.list),
                    ~ {
                      if (is.factor(.x[[1L]])) {
                        .x %>% purrr::map(as.character)
                      } else {
                        .x
                      }
                    }),
      # restore dates
      dplyr::across(any_of("date_time_created"),
                    untidy_date),
      # restore individual variables
      ## `files`
      dplyr::across(.cols = any_of("files"),
                    .fns = purrr::map,
                    .f = ~ {
                      
                      if ("date_time_attached" %in% colnames(.x)) {
                        
                        .x$date_time_attached %<>% untidy_date()
                      }
                      
                      .x %<>% rename_from_list(names_list = sub_v_names_files_inverse)
                    }),
      ## `inst_topics_excluded`
      dplyr::across(.cols = any_of("inst_topics_excluded"),
                    .fns = purrr::map,
                    .f = dplyr::recode,
                    "budgets" = "budget"),
      ## `inst_topics_only`
      dplyr::across(.cols = any_of("inst_topics_only"),
                    .fns = purrr::map,
                    .f = dplyr::recode,
                    "infrastructural acts" = "infrastructural act",
                    "competence shifts" = "competence shift",
                    "financial acts" = "financial act",
                    "financial acts (expenses)" = "financial act (expenses)",
                    "financial acts (taxes)" = "financial act (taxes)",
                    "financial acts (obligations)" = "financial act (obligations)",
                    "total revisions of the constitution" = "total revision of the constitution"),
      ## `inst_object_revision_divisibility`
      dplyr::across(any_of("inst_object_revision_divisibility"),
                    dplyr::recode,
                    "as a whole" = "Whole text only",
                    "split up" = "Splitting up possible",
                    "as variants" = "Variants possible",
                    "both split up and as variants" = "Variants / splitting up possible"),
      ## `inst_object_type`
      dplyr::across(any_of("inst_object_type"),
                    dplyr::recode,
                    "legal text (formulated proposal)" = "legal text (ausformulierter vorschlag)",
                    "legal text (general proposal)" = "legal text (allg. anregung)"),
      ## `inst_trigger_actor` / `inst_object_author`
      dplyr::across(any_of(c("inst_trigger_actor",
                             "inst_object_author")),
                    dplyr::recode,
                    "other institution" = "institution"),
      ## `inst_precondition_actor`
      dplyr::across(any_of("inst_precondition_actor"),
                    dplyr::recode,
                    "parliament and president" = "parliament and President",
                    "parliament and government" = "parliament and Government"),
      ## `level`
      dplyr::across(any_of("level"),
                    stringr::str_replace,
                    pattern = "sub-national",
                    replacement = "subnational"),
      ## `type`
      dplyr::across(any_of("type"),
                    dplyr::recode,
                    "citizens' assembly" = "citizen assembly",
                    .missing = "not provided"),
      ## `id`
      dplyr::across(any_of("id"),
                    ~ as.list(.x) %>% magrittr::set_names(rep("$oid", times = length(.)))),
      ## binary (dummies)
      dplyr::across(any_of("position_government"),
                    dplyr::recode,
                    "yes" = "Acceptance",
                    "no" = "Rejection"),
      dplyr::across(any_of("inst_has_urgent_legal_basis"),
                    dplyr::if_else,
                    true = "Urgent",
                    false = "Normal"),
      dplyr::across(any_of("inst_is_binding"),
                    dplyr::if_else,
                    true = "Binding",
                    false = "Non-binding"),
      dplyr::across(any_of("inst_is_counter_proposal"),
                    dplyr::if_else,
                    true = "Yes",
                    false = "No"),
      dplyr::across(any_of("inst_is_assembly"),
                    dplyr::if_else,
                    true = "Assembly",
                    false = "Ballot"),
      dplyr::across(any_of("inst_has_precondition"),
                    dplyr::if_else,
                    true = "Exists",
                    false = "Does not exist"),
      # uppercase first letter of various vars
      dplyr::across(.cols = any_of(c("level",
                                     "result",
                                     "type",
                                     "inst_legal_basis_type",
                                     "inst_object_type",
                                     "inst_object_legal_level",
                                     "inst_object_revision_extent",
                                     "inst_trigger_type",
                                     "inst_trigger_actor_level",
                                     "inst_trigger_time_limit",
                                     "inst_quorum_approval",
                                     "inst_precondition_decision",
                                     # vars containing uppercase abbreviations
                                     "inst_object_author",
                                     "inst_trigger_actor",
                                     "inst_precondition_actor")),
                    .fns = pal::sentenceify,
                    punctuation_mark = ""),
      dplyr::across(.cols = any_of(c("inst_object_revision_modes",
                                     "inst_topics_only",
                                     "inst_topics_excluded")),
                    .fns = purrr::map,
                    .f = pal::sentenceify,
                    punctuation_mark = ""),
      # restore NA values (always assumed to mean 'not provided', never 'unknown')
      dplyr::across(any_of("result"),
                    tidyr::replace_na,
                    replace = "Not provided"),
      dplyr::across(any_of(c("subterritories_yes", "subterritories_no")),
                    tidyr::replace_na,
                    replace = -2.0),
      dplyr::across(where(is.character),
                    tidyr::replace_na,
                    replace = ""),
      dplyr::across(where(is.integer),
                    tidyr::replace_na,
                    replace = -2L)
    ) %>%
    # restore variable names
    # TODO: find out why this (and only this) call of `rename_from_list()` causes the warning:
    #       "Unquoting language objects with `!!!` is deprecated as of rlang 0.4.0."
    rename_from_list(names_list = v_names_inverse)
  
  # restore `tags`
  tags_v_names <- paste0("tags_tier_", 1:3)
  tags_vx_present <- tags_v_names %in% colnames(data)
  
  if (any(tags_vx_present)) {
    
    if (!all(tags_vx_present)) {
      tags_vx_missing <- tags_v_names %>% setdiff(tags_vx_present)
      cli::cli_abort("There are only part of all {.var tags_tier_#} variables present. ",
                     "{cli::qty(tags_vx_missing)} The following variable{?s} are missing: {.var {tags_vx_missing}}")
    }
    
    data %<>%
      dplyr::mutate(tags = restore_tags(tags_tier_1,
                                        tags_tier_2,
                                        tags_tier_3)) %>%
      dplyr::select(-any_of(tags_v_names))
  }
  
  data %<>%
    # remove unknown columns
    dplyr::select(any_of(referendum_fields$all_flat)) %>%
    # remove variable labels
    labelled::remove_var_label()
  
  if (!as_tibble) {
    
    # remove nested field prefixes
    data %<>% dplyr::rename_with(.cols = matches("^(categories|context|title)\\."),
                                 .fn = ~ stringr::str_remove(string = .x,
                                                             pattern = "^\\w+?\\."))
    
    # restore nested structure
    categories_fields_present <-
      names(v_names) %>%
      stringr::str_subset(pattern = "^categories\\.") %>%
      stringr::str_remove(pattern = "^categories\\.") %>%
      intersect(colnames(data))
    
    context_fields_present <-
      names(v_names) %>%
      stringr::str_subset(pattern = "^context\\.") %>%
      stringr::str_remove(pattern = "^context\\.") %>%
      intersect(colnames(data))
    
    title_fields_present <-
      names(v_names) %>%
      stringr::str_subset(pattern = "^title\\.") %>%
      stringr::str_remove(pattern = "^title\\.") %>%
      intersect(colnames(data))
    
    if (length(categories_fields_present)) {
      data %<>% tidyr::nest(categories = all_of(categories_fields_present))
    }
    if (length(context_fields_present)) {
      data %<>% tidyr::nest(context = all_of(context_fields_present))
    }
    if (length(title_fields_present)) {
      data %<>% tidyr::nest(title = all_of(title_fields_present))
    }
    
    data %<>%
      # convert to list
      dplyr::group_split(dplyr::row_number(),
                         .keep = FALSE) %>%
      purrr::map(as.list) %>%
      # tweak list structure
      purrr::modify_depth(.depth = 1L,
                          .f = ~
                            .x %>%
                            # flatten unnecessarily nested elements
                            purrr::modify_at(.at = "tags",
                                             .f = flatten_array_as_is) %>%
                            # convert nested tibbles to lists
                            purrr::modify_at(.at = "files",
                                             .f = ~
                                               .x[[1L]] %>%
                                               purrr::when(is.null(.) ~ list(),
                                                           ~ dplyr::group_split(.tbl = .,
                                                                                dplyr::row_number(),
                                                                                .keep = FALSE) %>%
                                                             purrr::map(as.list))) %>%
                            purrr::modify_at(.at = c("archive",
                                                     "categories",
                                                     "context",
                                                     "title"),
                                             .f = ~
                                               .x %>%
                                               purrr::map(as.list) %>%
                                               unlist(recursive = FALSE)) %>%
                            # reduce nesting of nested tibble
                            purrr::when(is.null(purrr::pluck(., "context", "votes_per_canton")) ~ .,
                                        ~ purrr::modify_in(.x = .,
                                                           .where = c("context", "votes_per_canton"),
                                                           .f = dplyr::first)) %>%
                            # reduce nesting of array fields
                            purrr::when(is.null(purrr::pluck(., "categories", "action")) ~ .,
                                        ~ purrr::modify_in(.x = .,
                                                           .where = c("categories", "action"),
                                                           .f = flatten_array_as_is)) %>%
                            purrr::when(is.null(purrr::pluck(., "categories", "excluded_topics")) ~ .,
                                        ~ purrr::modify_in(.x = .,
                                                           .where = c("categories", "excluded_topics"),
                                                           .f = flatten_array_as_is)) %>%
                            purrr::when(is.null(purrr::pluck(., "categories", "special_topics")) ~ .,
                                        ~ purrr::modify_in(.x = .,
                                                           .where = c("categories", "special_topics"),
                                                           .f = flatten_array_as_is)))
  }
  
  data
}
```

### MongoDB query filter operator helpers

#### `$date`

Example query found in the mongolite documentation: <https://jeroen.github.io/mongolite/query-data.html#select-by-date>

```{r}
query_filter_date <- function(min,
                              max) {
  
  list(`$gte` = list(`$date` = lubridate::as_datetime(min)) %>% purrr::compact(),
       `$lte` = list(`$date` = lubridate::as_datetime(max)) %>% purrr::compact()) %>%
    purrr::compact()
}
```

#### `$in`

MongoDB documentation: <https://docs.mongodb.com/manual/reference/operator/query/in/#mongodb-query-op.-in>

```{r}
query_filter_in <- function(x) {
  
  x %>% purrr::when(is.null(.) || length(.) == 0L ~ NULL,
                    length(.) == 1 ~ .,
                    ~ list(`$in` = .))
}
```

### Tags-related

#### `restore_tags`

NOTE that this function doesn't restore *redundant* information (parent-tier tags). Once [issue \#41](https://github.com/ccmdesign/c2d-app/issues/41) is
resolved, we can deprecate this anyway.

```{r}
restore_tags <- function(tags_tier_1,
                         tags_tier_2,
                         tags_tier_3) {
  list(tags_tier_1,
       tags_tier_2,
       tags_tier_3) %>%
    purrr::pmap(~ {
      
      ..1 %>%
        unlist() %>%
        as.character() %>%
        checkmate::assert_character(any.missing = FALSE,
                                    max.len = 3L)
      ..2 %>%
        unlist() %>%
        as.character() %>%
        checkmate::assert_character(any.missing = FALSE,
                                    max.len = 3L)
      ..3 %>%
        unlist() %>%
        as.character() %>%
        checkmate::assert_character(any.missing = FALSE,
                                    max.len = 3L,
                                    .var.name = "tags_tier_3")
      
      tags_hierarchy <- hierarchize_tags(tibble::tibble(tags_tier_1 = list(..1),
                                                        tags_tier_2 = list(..2),
                                                        tags_tier_3 = list(..3)))
      
      tags <- tags_hierarchy$tag_tier_3 %>% setdiff(NA_character_)
      
      if (length(tags) < 3L) {
        
        tags <-
          tags_hierarchy %>%
          dplyr::filter(is.na(tag_tier_3)) %$%
          tag_tier_2 %>%
          setdiff(NA_character_) %>%
          c(tags)
      }
      
      if (length(tags) < 3L) {
        
        tags <-
          tags_hierarchy %>%
          dplyr::filter(is.na(tag_tier_3) & is.na(tag_tier_2)) %$%
          tag_tier_1 %>%
          setdiff(NA_character_) %>%
          c(tags)
      }
      
      tags
    })
}
```

#### `tags`

```{r}
tags <- function(tiers = 1:3) {
  
  checkmate::assert_integerish(tiers,
                               lower = 1L,
                               upper = 3L,
                               any.missing = FALSE,
                               unique = TRUE)
  tag_set <- character()
  
  if (1L %in% tiers) {
    tag_set %<>% c(data_tags$tag_tier_1)
  }
  
  if (2L %in% tiers) {
    tag_set %<>% c(data_tags$tag_tier_2)
  }
  
  if (3L %in% tiers) {
    tag_set %<>% c(data_tags$tag_tier_3)
  }
  
  tag_set %>% setdiff(NA_character_) %>% unique()
}
```

## Constants

```{r}
pkg <- utils::packageName()
```

### C2D referendum database fields

Field names as returned by the `referendums` API endpoint.

```{r}
referendum_fields <- list()
```

#### All

```{r}
referendum_fields$all <- c("_id",
                           "archive",
                           "canton",
                           "categories",
                           "citizens_abroad",
                           "committee_name",
                           "context",
                           "country_code",
                           "country_name",
                           "created_on",
                           "date",
                           "draft",
                           "files",
                           "institution",
                           "level",
                           "municipality",
                           "number",
                           "remarks",
                           "result",
                           "tags",
                           "title",
                           "total_electorate",
                           "votes_empty",
                           "votes_invalid",
                           "votes_no",
                           "votes_yes")

referendum_fields$all_flat <-
  referendum_fields$all %>%
  setdiff(c("categories", "context", "title")) %>%
  union(c("categories.action",
          "categories.author_of_the_vote_object",
          "categories.counter_proposal",
          "categories.decision_quorum",
          "categories.degree_of_revision",
          "categories.excluded_topics",
          "categories.hierarchy_of_the_legal_norm",
          "categories.institutional_precondition",
          "categories.institutional_precondition_decision",
          "categories.institutional_precondition_decision_actor",
          "categories.legal_act_type",
          "categories.official_status",
          "categories.referendum_text_options",
          "categories.special_topics",
          "categories.turnout_quorum",
          "categories.vote_object",
          "categories.vote_result_status",
          "categories.vote_trigger",
          "categories.vote_trigger_actor",
          "categories.vote_trigger_number",
          "categories.vote_trigger_state_level",
          "categories.vote_trigger_time",
          "categories.vote_venue",
          "context.national_council_abstentions",
          "context.national_council_no",
          "context.national_council_yes",
          "context.recommendation",
          "context.states_council_abstentions",
          "context.states_council_no",
          "context.states_council_yes",
          "context.states_no",
          "context.states_yes",
          "context.votes_per_canton",
          "title.de",
          "title.en",
          "title.fr"))
```

#### Mandatory

The following fields are mandatory when adding/editing referendums via the [admin portal](https://admin.c2d.ch/), but not *strictly* mandatory schema-wise:

| field             | remarks                                                                                                               |
|-------------------|-----------------------------------------------------------------------------------------------------------------------|
| `institution`     | mandatory by the API for new entries, but still contains `NA`s from before relaunch                                   |
| `title.en`        | mandatory by the API for new entries, but still contains `NA`s from before relaunch                                   |
| `citizens_abroad` | not mandatory by the API, but omitting it [leads to serious problems](https://github.com/ccmdesign/c2d-app/issues/50) |
| `votes_no`        | not mandatory by the API, but omitting it [leads to serious problems](https://github.com/ccmdesign/c2d-app/issues/50) |
| `votes_yes`       | not mandatory by the API, but omitting it [leads to serious problems](https://github.com/ccmdesign/c2d-app/issues/50) |

```{r}
referendum_fields$required <- c("_id",
                                "country_code",
                                "country_name",
                                "created_on",
                                "level",
                                "total_electorate")
```

### Messages

```{r}
mime_error_suffix <- "This indicates either some network issue or a change in the C2D API."
```

### Variable renamings

This maps the field names of the back-end's MongoDB to the final variable names as found in the [C2D codebook](https://rpkg.dev/c2d/articles/codebook.html). The
renamings should eventually all be upstreamed and this mapping shall be become superfluous.

```{r}
                # old name                                               new name
v_names <- list("_id"                                                  = "id",
                "number"                                               = "id_sudd",
                "canton"                                               = "subnational_entity_name",
                "title.de"                                             = "title_de",
                "title.en"                                             = "title_en",
                "title.fr"                                             = "title_fr",
                "context.states_no"                                    = "subterritories_no",
                "context.states_yes"                                   = "subterritories_yes",
                "total_electorate"                                     = "electorate_total",
                "citizens_abroad"                                      = "electorate_abroad",
                "context.votes_per_canton"                             = "votes_per_subterritory",
                "context.national_council_yes"                         = "lower_house_yes",
                "context.national_council_no"                          = "lower_house_no",
                "context.national_council_abstentions"                 = "lower_house_abstentions",
                "context.states_council_yes"                           = "upper_house_yes",
                "context.states_council_no"                            = "upper_house_no",
                "context.states_council_abstentions"                   = "upper_house_abstentions",
                "context.recommendation"                               = "position_government",
                "draft"                                                = "is_draft",
                "created_on"                                           = "date_time_created",
                "institution"                                          = "type",
                "categories.official_status"                           = "inst_legal_basis_type",
                "categories.legal_act_type"                            = "inst_has_urgent_legal_basis",
                "categories.vote_result_status"                        = "inst_is_binding",
                "categories.counter_proposal"                          = "inst_is_counter_proposal",
                "categories.vote_venue"                                = "inst_is_assembly",
                "categories.vote_trigger"                              = "inst_trigger_type",
                "categories.vote_trigger_actor"                        = "inst_trigger_actor",
                "categories.vote_trigger_state_level"                  = "inst_trigger_actor_level",
                "categories.vote_trigger_number"                       = "inst_trigger_threshold",
                "categories.vote_trigger_time"                         = "inst_trigger_time_limit",
                "categories.vote_object"                               = "inst_object_type",
                "categories.author_of_the_vote_object"                 = "inst_object_author",
                "categories.hierarchy_of_the_legal_norm"               = "inst_object_legal_level",
                "categories.degree_of_revision"                        = "inst_object_revision_extent",
                "categories.action"                                    = "inst_object_revision_modes",
                "categories.referendum_text_options"                   = "inst_object_revision_divisibility",
                "categories.turnout_quorum"                            = "inst_quorum_turnout",
                "categories.decision_quorum"                           = "inst_quorum_approval",
                "categories.institutional_precondition"                = "inst_has_precondition",
                "categories.institutional_precondition_decision_actor" = "inst_precondition_actor",
                "categories.institutional_precondition_decision"       = "inst_precondition_decision",
                "categories.special_topics"                            = "inst_topics_only",
                "categories.excluded_topics"                           = "inst_topics_excluded")

                                 # old name     new name
sub_v_names <- list(files = list("date"       = "date_time_attached",
                                 "object_key" = "s3_object_key",
                                 "size"       = "file_size",
                                 "deleted"    = "is_deleted"))
```

## Data

The R code chunks below are *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) (`purl = FALSE`). The chunks
have to be executed at least once to generate the necessary internal data file [`R/sysdata.rda`](https://r-pkgs.org/data.html#data-sysdata).

### Codebook data

TODO: Once the toml\* package is finished, use it instead to do the stuff below more generically.

```{r, purl = FALSE}
data_codebook <- {
  
  pal::assert_pkg("assertr")
  codebook <- read_toml("data-raw/c2d_codebook.toml")
  
  c(codebook$`01_main`$item,
    codebook$`02_institutional`$`01_status`$item,
    codebook$`02_institutional`$`02_trigger`$item,
    codebook$`02_institutional`$`03_object`$item,
    codebook$`02_institutional`$`04_other`$item) %>%
    purrr::map_dfr(~ tibble::tibble(variable_name = .x$variable_name %||% NA_character_,
                                    variable_label = .x$variable_label %||% NA_character_,
                                    is_multi_valued = .x$is_multi_valued %||% FALSE,
                                    applicability_constraint = .x$applicability_constraint %||% NA_character_,
                                    variable_values = list(.x$variable_values) %||% list(character()),
                                    value_label_prefix = .x$value_label_prefix %||% NA_character_,
                                    value_labels = list(.x$value_labels) %||% list(character()),
                                    value_label_suffix = .x$value_label_suffix %||% NA_character_,
                                    value_default = list(.x$value_default) %||% list(character()),
                                    value_scale = .x$value_scale %||% NA_character_)) %>%
    # integrity checks
    assertr::assert(predicate = assertr::not_na,
                    variable_name,
                    variable_label,
                    value_scale,
                    is_multi_valued)
}
```

### Tags

```{r, purl = FALSE}
data_tags <-
  # download file from private CCM Design repo
  pal::gh_text_file(owner = "ccmdesign",
                    name = "c2d-app",
                    rev = "master",
                    path = "ch.c2d/web/themes.json") %>%
  jsonlite::fromJSON(flatten = FALSE) %>%
  tibble::as_tibble() %>%
  # create tidy data version
  dplyr::rename(tag_tier_1 = name,
                tag_tier_2 = children) %>%
  tidyr::unnest_longer(col = tag_tier_2) %$%
  dplyr::bind_cols(tag_tier_1 = tag_tier_1,
                   tag_tier_2) %>%
  dplyr::rename(tag_tier_2 = name,
                tag_tier_3 = children) %>%
  # ensure consistent col subtypes
  dplyr::mutate(tag_tier_3 = tag_tier_3 %>% purrr::map(~ { if (is.list(.x)) character() else .x })) %>%
  tidyr::unnest_longer(col = tag_tier_3)
```

### Constants pre-computed for performance purposes

```{r, purl = FALSE}
# NOTE: `var_lbls` is assigned here because when inlined, it somehow breaks `devtools::run_examples()` during pkg check ("C stack usage (...) is too close to
#       the limit"); maybe too large fn stack? (modifying `pal::strip_md()` to minimize the fn stack was tried but didn't help)
var_lbls <-
  data_codebook$variable_label %>%
  pal::strip_md() %>%
  as.list() %>%
  magrittr::set_names(value = data_codebook$variable_name)
```

Finally, we have to save all the external data objects to `data/*.RData` as well as the internal data objects to `R/sysdata.rda`.

Note that `data_codebook` is saved as internal data and later exported because external data doesn't support `@aliases`.

```{r, purl = FALSE}
usethis::use_data(data_codebook,
                  data_tags,
                  var_lbls,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```

# Referendum data

## `referendums`

```{r}
#' Get referendum data
#'
#' Downloads the referendum data from the C2D Database. See the [`codebook`][codebook] for a detailed description of all variables.
#'
#' @param use_cache `r pkgsnip::param_label("use_cache")`
#' @param cache_lifespan `r pkgsnip::param_label("cache_lifespan")`
#' @param incl_archive Whether or not to include an `archive` column containing data from an earlier, obsolete state of the C2D database.
#' @inheritParams assemble_query_filter
#' @inheritParams tidy_referendums
#'
#' @return `r pkgsnip::return_label("data")`
#' @family referendum
#' @export
#'
#' @examples
#' \dontrun{
#' # get all referendums (excl. drafts)
#' c2d::referendums()}
#' 
#' # get only referendums in Austria and Australia on subnational level
#' c2d::referendums(country_code = c("AT", "AU"),
#'                  level = "subnational")
#' 
#' # provide custom `query_filter` for more complex queries like regex matches
#' # cf. https://docs.mongodb.com/manual/reference/operator/query/regex/
#' c2d::referendums(query_filter = '{"country_code":{"$regex":"A."}}')
referendums <- function(use_cache = TRUE,
                        cache_lifespan = "1 week",
                        incl_archive = FALSE,
                        is_draft = FALSE,
                        country_code = NULL,
                        subnational_entity_name = NULL,
                        municipality = NULL,
                        level = NULL,
                        type = NULL,
                        date = NULL,
                        date_time_created_min = NULL,
                        date_time_created_max = NULL,
                        query_filter = NULL,
                        tidy = TRUE) {
  
  pkgpins::with_cache(.fn = ~ {
    
    checkmate::assert_flag(incl_archive)
    
    # retrieve data
    data <-
      httr::RETRY(verb = "GET",
                  url = "https://services.c2d.ch/referendums",
                  query = list(mode = "stream",
                               format = "json",
                               filter = assemble_query_filter(country_code = country_code,
                                                              subnational_entity_name = subnational_entity_name,
                                                              municipality = municipality,
                                                              level = level,
                                                              type = type,
                                                              date = date,
                                                              is_draft = is_draft,
                                                              date_time_created_min = date_time_created_min,
                                                              date_time_created_max = date_time_created_max,
                                                              query_filter = query_filter)),
                  times = 5L) %>%
      # ensure we actually got a JSON response
      pal::assert_mime_type(mime_type = "application/json",
                            msg_suffix = mime_error_suffix) %>%
      # extract JSON
      httr::content(as = "text",
                    encoding = "UTF-8") %>%
      # convert JSON to list
      # NOTE that we cannot rely on params `simplify*` and `flatten` because of varying field lengths in API result (depending on `query`)
      jsonlite::fromJSON(simplifyVector = FALSE,
                         simplifyDataFrame = FALSE,
                         simplifyMatrix = FALSE,
                         flatten = FALSE) %$%
      items %>%
      # tidy data
      tidy_referendums(tidy = tidy)
    
    # exclude `archive` if requested
    if (!incl_archive) data %<>% dplyr::select(-any_of("archive"))
    
    # return data
    data
  },
  .use_cache = use_cache,
  .cache_lifespan = cache_lifespan,
  .pkg = pkg)
}
```

## `referendum`

```{r}
#' Get a single referendum's data
#'
#' Downloads a single referendum's data from the C2D Database. See the [`codebook`][codebook] for a detailed description of all variables.
#'
#' @param id The referendum's unique [identifier](https://rpkg.dev/c2d/articles/codebook.html#id).
#' @inheritParams referendums
#'
#' @inherit referendums return
#' @family referendum
#' @export
#'
#' @examples
#' c2d::referendum(id = "5bbbe26a92a21351232dd73f")
referendum <- function(id,
                       incl_archive = FALSE,
                       tidy = TRUE) {
  
  checkmate::assert_string(id,
                           min.chars = 1L)
  checkmate::assert_flag(incl_archive)
  checkmate::assert_flag(tidy)
  
  # retrieve data
  data <-
    httr::RETRY(verb = "GET",
                url = paste0("https://services.c2d.ch/referendums/", id),
                times = 5L) %>%
    # ensure we actually got a JSON response
    pal::assert_mime_type(mime_type = "application/json",
                          msg_suffix = mime_error_suffix) %>%
    # extract JSON
    httr::content(as = "text",
                  encoding = "UTF-8") %>%
    # convert JSON to list
    # NOTE that we cannot rely on params `simplify*` and `flatten` because of varying field lengths in API result
    jsonlite::fromJSON(simplifyVector = FALSE,
                       simplifyDataFrame = FALSE,
                       simplifyMatrix = FALSE,
                       flatten = FALSE) %>%
    # tidy data
    tidy_referendums(tidy = tidy)
  
  # exclude `archive` if requested
  if (!incl_archive) data %<>% dplyr::select(-any_of("archive"))
  
  # return data
  data
}
```

## `download_file_attachment`

```{r}
#' Download file attachment
#'
#' Downloads a file attachment from the C2D database. The necessary `s3_object_key`s identifying individual files are found in the `files` list column returned
#' by [referendums()].
#'
#' @param s3_object_key The key uniquely identifying the file in the C2D [Amazon S3 bucket](https://en.wikipedia.org/wiki/Amazon_S3#Design). A character scalar.
#' @param path The path where the downloaded file is written to. If a directory, the original filename returned by the C2D services API will be used.
#'
#' @return A [response object][httr::response], invisibly.
#' @family referendum
#' @export
#'
#' @examples
#' \dontrun{
#' # get all file object keys...
#' c2d::referendums()$files |>
#'   purrr::map_depth(2L, purrr::pluck, "s3_object_key") |>
#'   purrr::flatten() |>
#'   purrr::flatten_chr() |>
#'   # ...select first three keys...
#'   magrittr::extract(1:3) |>
#'   # ...and download the corresponding files to the current working dir
#'   purrr::walk(c2d::download_file_attachment)}
download_file_attachment <- function(s3_object_key,
                                     path = ".") {
  
  checkmate::assert_string(s3_object_key)
  checkmate::assert_atomic(path)
  
  if (fs::dir_exists(path)) {
    checkmate::assert_directory(path,
                                access = "rw")
    use_original_filename <- TRUE
  } else {
    checkmate::assert_path_for_output(path,
                                      overwrite = TRUE)
    use_original_filename <- FALSE
  }
  
  temp_path <- fs::file_temp()
  
  response <- httr::RETRY(verb = "GET",
                          url = paste0("https://services.c2d.ch/s3_objects/", s3_object_key),
                          httr::write_disk(path = temp_path),
                          times = 5L)
  
  if (use_original_filename) {
    
    final_path <-
      response %>%
      httr::headers() %$%
      `content-disposition` %>%
      stringr::str_extract(pattern = "(?<=filename=\").+?(?=\")") %>%
      fs::path(path, .)
    
  } else {
    final_path <- path
  }
  
  fs::file_move(path = temp_path,
                new_path = final_path)
  
  invisible(response)
}
```

## `add_referendums`

**Notes:**

-   For additions, the following JSON fields are mandatory:

    | field                    | notes                                                                                                             |
    |--------------------------|-------------------------------------------------------------------------------------------------------------------|
    | `country_code`           | must be a valid choice                                                                                            |
    | `canton`                 | only mandatory if `"level":"Sub-national"`                                                                        |
    | `municipality`           | only mandatory if `"level":"Local"`                                                                               |
    | `level`                  | must be a valid choice                                                                                            |
    | `date`                   | must be a valid date (ISO 8601 date string)                                                                       |
    | `institution`            | must be a valid choice                                                                                            |
    | `title.en`               | must contain at least one non-whitespace character                                                                |
    | `result`                 | must be a valid choice ("Yes", "No" or "Unknown")                                                                 |
    | `context.recommendation` | only mandatory if `"country_code":"CH","level":"National"`; must be a valid choice                                |
    | `total_electorate`       | must be a valid integer                                                                                           |
    | `citizens_abroad`        | not actually mandatory, but [serious problems arise when omitted](https://github.com/ccmdesign/c2d-app/issues/50) |
    | `votes_yes`              | not actually mandatory, but [serious problems arise when omitted](https://github.com/ccmdesign/c2d-app/issues/50) |
    | `votes_no`               | not actually mandatory, but [serious problems arise when omitted](https://github.com/ccmdesign/c2d-app/issues/50) |
    | `votes_empty`            | not actually mandatory, but [serious problems arise when omitted](https://github.com/ccmdesign/c2d-app/issues/50) |
    | `votes_invalid`          | not actually mandatory, but [serious problems arise when omitted](https://github.com/ccmdesign/c2d-app/issues/50) |

-   If `draft` is not provided, it defaults to `true`.

-   `country_name` is ignored and automatically derived from `country_code`.

-   Nested fields like `categories` can be partially provided.

-   The `context.*` fields are ignored unless `"country_code":"CH","level":"National"`, cf. [issue \#52](https://github.com/ccmdesign/c2d-app/issues/52).

-   The order of the JSON fields doesn't matter.

-   The API performs input validation on certain fields, and if it fails, it responds with a message like:

    ``` {.json}
    {"error": {"id": "validation_failed", "value": {"categories.author_of_the_vote_object": "invalid"}}}
    ```

-   A minimal sample cURL request to add a test entry looks as follows:

    ``` {.curl}
    curl 'https://services.c2d.ch/referendums' \
         --request POST \
         --header 'Content-Type: application/json' \
         --header 'Authorization: Bearer eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJzdWIiOiI1YzhiYTE3ZDYxYWY4NDYyMmRkZThjNjEiLCJleHAiOjE2MzAxODY0ODB9.lpKWUPyYh7yGWN1AKBWkjhrsQIVeIH71DUOA4hpJhR8' \
         --data-raw $'{"title":{"en":"test test test"},"date":"2222-02-22","level":"National","country_code":"AT","institution":"Citizens\' initiative","votes_no":-1,"votes_yes":-1,"votes_empty":-1,"votes_invalid":-1,"total_electorate":-1,"citizens_abroad":-1,"result":"Unknown"}'
    ```

```{r}
#' Add new referendums to the C2D database
#'
#' Adds new referendum entries to the C2D database via [its API](https://github.com/ccmdesign/c2d-app/blob/master/docs/services.md#3-referendum-routes).
#'
#' @details
#' Note that the following variables are not supported, i.e. simply dropped from `data`: `files`
#'
#' @param data The new referendum data. A [tibble][tibble::tbl_df] that must contain an [`id`](https://rpkg.dev/c2d/articles/codebook.html#id) column 
#'   identifying the referendums plus additional [valid columns][codebook] containing the values for the corresponding database fields.
#' @param email The e-mail address of the C2D API user account to be used for authentication. A character scalar.
#' @param password The password of the C2D API user account to be used for authentication. A character scalar.
#'
#' @return `data`, invisibly.
#' @family referendum
#' @export
add_referendums <- function(data,
                            email = names(getOption("c2d.credentials")[1L]),
                            password = getOption("c2d.credentials")[1L]) {
  # ensure `data` is valid
  checkmate::assert_data_frame(data,
                               min.rows = 1L)
  
  ## ensure forbidden columns are absent
  if ("id" %in% colnames(data)) {
    cli::cli_abort(paste0("{.arg data} mustn't contain an {.var id} column. It is automatically set by the C2D API back-end. Did you mean to {.fun ",
                          "edit_referendums} instead?"))
  }
  if ("created_on" %in% colnames(data)) {
    cli::cli_abort("{.arg data} mustn't contain a {.var created_on} column. This date is automatically set by the C2D API back-end.")
  }
  if ("country_name" %in% colnames(data)) {
    cli::cli_abort("{.arg data} mustn't contain a {.var country_name} column. It is automatically set by the C2D API back-end based on {.var country_code}.")
  }
  
  ## ensure mandatory columns are present
  c("country_code",
    "level",
    "date",
    "title_en",
    "result",
    "electorate_total",
    "electorate_abroad",
    "votes_yes",
    "votes_no",
    "votes_empty",
    "votes_invalid",
    "is_draft",
    "type") %>%
    purrr::walk(~ if (!(.x %in% colnames(data))) cli::cli_abort(paste0("Mandatory column {.var ", .x, "} is missing from {.arg data}.")))
  
  ## ensure columns are valid
  assert_cols_valid(data)
  
  # drop disabled variables
  data %<>% drop_disabled_vx("files")
  
  # convert data to MongoDB schema
  json_items <-
    data %>%
    # restore MongoDB fields
    untidy_referendums() %>%
    # convert to JSON
    purrr::map(jsonlite::toJSON,
               auto_unbox = TRUE,
               digits = NA)
  
  # add data to the MongoDB via HTTP POST on `/referendums` API endpoint
  responses <-
    json_items %>%
    purrr::map(~ httr::RETRY(verb = "POST",
                             url = "https://services.c2d.ch/referendums",
                             config = httr::add_headers(Authorization = paste("Bearer", auth_session(email = email,
                                                                                                     password = password)),
                                                        Connection = "keep-alive"),
                             body = .x,
                             times = 3L,
                             httr::content_type_json()) %>%
                 # ensure we actually got a JSON response
                 pal::assert_mime_type(mime_type = "application/json",
                                       msg_suffix = mime_error_suffix) %>%
                 # extract JSON string
                 httr::content(as = "text",
                               encoding = "UTF-8"))
  
  # throw warnings for unsuccessful API calls
  purrr::walk2(.x = seq_along(responses),
               .y = responses,
               .f = ~ if (!isTRUE(jsonlite::fromJSON(.y)$ok)) {
                 cli::cli_warn("Failed to add the {.x}. referendum. The API server responded with error {.field {.y}}.")
               })
  
  invisible(data)
}
```

## `edit_referendums`

**Notes:**

-   For edits, the only mandatory JSON field is `draft`. Only the fields that are supposed to change need to be supplied. Nested fields like `categories` can be
    partially overwritten, too.

    `country_name` is ignored and automatically derived from `country_code`.

-   The `context.*` fields are ignored unless `"country_code":"CH","level":"National"`, cf. [issue \#52](https://github.com/ccmdesign/c2d-app/issues/52).

-   The order of the JSON fields doesn't matter.

-   The API performs input validation on certain fields, and if it fails, it responds with a message like:

    ``` {.json}
    {"error": {"id": "validation_failed", "value": {"categories.author_of_the_vote_object": "invalid"}}}
    ```

-   A minimal sample cURL request to change a test entry looks as follows:

    ``` {.sh}
    curl 'https://services.c2d.ch/referendums/610005e2c72633da60229938' \
         --request PUT \
         --header 'Content-Type: application/json' \
         --header 'Authorization: Bearer eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJzdWIiOiI1YzhiYTE3ZDYxYWY4NDYyMmRkZThjNjEiLCJleHAiOjE2MzAxODY0ODB9.lpKWUPyYh7yGWN1AKBWkjhrsQIVeIH71DUOA4hpJhR8' \
         --header 'Connection: keep-alive' \
         --data-raw $'{"draft":true,"title":{"de":"salims cat test"}}'
    ```

```{r}
#' Edit existing referendums in the C2D database
#'
#' Edits existing referendum entries in the C2D database via [its API](https://github.com/ccmdesign/c2d-app/blob/master/docs/services.md#3-referendum-routes).
#'
#' @inherit add_referendums details
#' @param data The updated referendum data. A [tibble][tibble::tbl_df] that must contain an [`id`](https://rpkg.dev/c2d/articles/codebook.html#id) column 
#'   identifying the referendums plus a [valid column][codebook] containing the new values for each database field that should be updated.
#' @inheritParams add_referendums
#'
#' @inherit add_referendums return
#' @family referendum
#' @export
edit_referendums <- function(data,
                             email = names(getOption("c2d.credentials")[1L]),
                             password = getOption("c2d.credentials")[1L]) {
  # ensure `data` is valid
  checkmate::assert_data_frame(data,
                               min.rows = 1L)
  
  ## ensure forbidden columns are absent
  if ("created_on" %in% colnames(data)) {
    cli::cli_abort(paste0("{.arg data} mustn't contain a {.var created_on} column. This date is automatically set by the C2D API back-end and not supposed to ",
                          "be changed."))
  }
  if ("country_name" %in% colnames(data)) {
    cli::cli_abort("{.arg data} mustn't contain a {.var country_name} column. It is automatically set by the C2D API back-end based on {.var country_code}.")
  }
  
  ## ensure mandatory columns are present
  c("id",
    "is_draft") %>%
    purrr::walk(~ if (!(.x %in% colnames(data))) cli::cli_abort(paste0("Mandatory column {.var ", .x, "} is missing from {.arg data}.")))
  
  ## ensure columns are valid
  assert_cols_valid(data)
  
  # drop disabled variables
  data %<>% drop_disabled_vx("files")
  
  # convert data to MongoDB schema
  ids <- data$id
  
  json_items <-
    data %>%
    # drop `id`
    dplyr::select(-id) %>%
    # restore MongoDB fields
    untidy_referendums() %>%
    # convert to JSON
    purrr::map(jsonlite::toJSON,
               auto_unbox = TRUE,
               digits = NA)
  
  # edit data in the MongoDB via HTTP PUT on `/referendums` API endpoint
  responses <- purrr::map2(.x = ids,
                           .y = json_items,
                           .f = ~
                             httr::RETRY(verb = "PUT",
                                         url = paste0("https://services.c2d.ch/referendums/",.x),
                                         config = httr::add_headers(Authorization = paste("Bearer", auth_session(email = email,
                                                                                                                 password = password)),
                                                                    Connection = "keep-alive"),
                                         body = .y,
                                         times = 3L,
                                         httr::content_type_json()) %>%
                             # ensure we actually got a JSON response
                             pal::assert_mime_type(mime_type = "application/json",
                                                   msg_suffix = mime_error_suffix) %>%
                             # extract JSON string
                             httr::content(as = "text",
                                           encoding = "UTF-8"))
  
  # throw warnings for unsuccessful API calls
  purrr::walk2(.x = ids,
               .y = responses,
               .f = ~ if (!isTRUE(jsonlite::fromJSON(.y)$ok)) {
                 cli::cli_warn("Failed to edit referendum with {.var id} {.val {.x}}. The API server responded with {.field {.y}}.")
               })
  
  invisible(data)
}
```

## `validate_referendums`

This function performs various data validation steps to ensure there are no errors in the database.

```{r}
#' Validate referendums
#'
#' @param data The referendum data to validate, as returned by [referendums()].
#' @param check_applicability_constraint Whether or not to check that no applicability constraints as defined in the [codebook][data_codebook] are violated.
#' @param check_id_sudd_prefix Whether or not to check that all [`id_sudd`](https://rpkg.dev/c2d/articles/codebook.html#id-sudd) prefixes are valid.
#'
#' @return `data`, invisibly.
#' @family referendum
#' @export
validate_referendums <- function(data,
                                 check_applicability_constraint = TRUE,
                                 check_id_sudd_prefix = TRUE) {
  
  checkmate::assert_data_frame(data,
                               min.rows = 1L)
  checkmate::assert_subset(colnames(data),
                           choices = data_codebook$variable_name)
  checkmate::assert_flag(check_applicability_constraint)
  checkmate::assert_flag(check_id_sudd_prefix)
  
  # check columns
  assert_cols_valid(data = data,
                    action = cli::cli_alert_warning)
  
  # check applicability constraints
  if (check_applicability_constraint) {
    
    v_names_violated <-
      data_codebook %>%
      dplyr::filter(variable_name %in% colnames(data)
                    & !is.na(applicability_constraint)) %$%
      purrr::map2_lgl(.x = magrittr::set_names(x = variable_name,
                                               value = variable_name),
                      .y = applicability_constraint,
                      .f = ~ {
                        
                        data %>%
                          dplyr::filter(!eval(parse(text = .y))) %$%
                          eval(as.symbol(.x)) %>%
                          {is.na(.) | purrr::map_lgl(., is.null)} %>%
                          all()
                      }) %>%
    magrittr::extract(. %in% FALSE) %>%
    names()
    
    n_v_names_violated <- length(v_names_violated)
    
    if (n_v_names_violated) {
      
      cli::cli_alert_warning("Applicability constraints are violated for {n_v_names_violated} variable{?s}:")
      
      paste0("{.var ", v_names_violated, "}") %>%
        magrittr::set_names(rep("x",
                                times = length(.))) %>%
        cli::cli_bullets()
      
      first_v_name_violated <- v_names_violated[1L]
      
      cli::cli_text("To get the applicability constraint of e.g. {.var {first_v_name_violated}}, run:")
      cat("\n")
      cli::cli_code(c("data_codebook %>%",
                      glue::glue("  dplyr::filter(variable_name == \"{first_v_name_violated}\") %$%"),
                      "  applicability_constraint"))
      cat("\n")
      cli::cli_text("To inspect the entries in violation of the above applicability constraint, run:")
      cat("\n")
      cli::cli_code(c("data_codebook %>%",
                      glue::glue("  dplyr::filter(variable_name == \"{first_v_name_violated}\") %$%"),
                      "  dplyr::filter(.data = data,",
                      "                !eval(parse(text = applicability_constraint))) %$%",
                      glue::glue("  {first_v_name_violated}")))
    }
  }
  
  # check `id_sudd` prefix if requested
  if (check_id_sudd_prefix) {
    
    if (!all(c("country_code", "id_sudd") %in% colnames(data))) {
      cli::cli_abort("Columns {.var country_code} and {.var id_sudd} must be present in {.arg data}.")
    }
    
    # define allowed exceptions
    allowed_exceptions <- tibble::tribble(
      ~country_code, ~id_sudd_prefix,
      # Curaao
      "CW", "an",
      # Szeklerland, cf. https://sudd.ch/event.php?id=hu042008
      "RO", "hu"
    )
    
    # assemble target country codes
    country_codes <- data$country_code %>% as.list()
    
    for (country_code in allowed_exceptions$country_code) {
      
      additional_country_codes <-
        allowed_exceptions %>%
        dplyr::filter(country_code == !!country_code) %$%
        id_sudd_prefix %>%
        stringr::str_to_upper()
      i_country_codes <- country_codes %>% purrr::map_lgl(~ country_code == .x) %>% which()
      
      for (i in i_country_codes) {
        country_codes[[i]] <- unique(c(country_codes[[i]], additional_country_codes))
      }
    }
    
    # add dummy indicating if target country codes match
    # TODO: instead of modifying input data, print cli msg with all relevant info!
    data$matches_id_sudd_prefix <-
      data$id_sudd %>%
      stringr::str_extract(pattern = "^..") %>%
      stringr::str_to_upper() %>%
      purrr::map2_lgl(.y = country_codes,
                      .f = ~ .x %in% .y)
    
    data$matches_id_sudd_prefix[is.na(data$id_sudd)] <- NA
  }
  
  invisible(data)
}
```

## `count_referendums`

```{r}
#' Count number of referendums
#'
#' Counts the number of referendums per [`level`](https://rpkg.dev/c2d/articles/codebook.html#level) in the C2D Database.
#'
#' @inheritParams assemble_query_filter
#'
#' @return A named list with `level` as names and referendum counts as values.
#' @family referendum
#' @export
#'
#' @examples
#' # the whole database (excl. drafts)
#' c2d::count_referendums()
#' 
#' # only Swiss and Austrian referendums
#' c2d::count_referendums(country_code = c("CH", "AT"))
#' 
#' # only Swiss referendums created between 2020 and 2021
#' c2d::count_referendums(country_code = "CH",
#'                        date_time_created_min = "2020-01-01",
#'                        date_time_created_max = "2021-01-01")
count_referendums <- function(is_draft = FALSE,
                              country_code = NULL,
                              subnational_entity_name = NULL,
                              municipality = NULL,
                              level = NULL,
                              type = NULL,
                              date = NULL,
                              date_time_created_min = NULL,
                              date_time_created_max = NULL,
                              query_filter = NULL) {
  httr::RETRY(verb = "GET",
              url = "https://services.c2d.ch/referendums/stats",
              query = list(filter = assemble_query_filter(country_code = country_code,
                                                          subnational_entity_name = subnational_entity_name,
                                                          municipality = municipality,
                                                          level = level,
                                                          type = type,
                                                          date = date,
                                                          is_draft = is_draft,
                                                          date_time_created_min = date_time_created_min,
                                                          date_time_created_max = date_time_created_max,
                                                          query_filter = query_filter)),
              times = 5L) %>%
    # ensure we actually got a JSON response
    pal::assert_mime_type(mime_type = "application/json",
                          msg_suffix = mime_error_suffix) %>%
    # parse response
    httr::content(as = "parsed") %$%
    votes %>%
    magrittr::set_names(names(.) %>% dplyr::recode("sub_national" = "subnational"))
}
```

## `search_referendums`

Currently, the C2D API [only supports searching in the English referendum
title](https://github.com/ccmdesign/c2d-app/blob/master/docs/services.md#3-referendum-routes).

```{r}
#' Search in English referendum titles
#'
#' Allows to use the C2D API's primitive search functionality. The search is not case-sensitive and no [fuzzy
#' search](https://en.wikipedia.org/wiki/Approximate_string_matching) is performed (i.e. only exact matches are returned).
#'
#' Note that this function is probably not of much use since it doesn't return any additional information about the matched referendums but only the English
#' titles.
#'
#' @param term The Search term. A character scalar.
#'
#' @return A character vector of English referendum titles matching the search `term`.
#' @family referendum
#' @export
#'
#' @examples
#' c2d::search_referendums("freedom")
search_referendums <- function(term) {
  
  httr::RETRY(verb = "GET",
              url = "https://services.c2d.ch/referendums",
              query = list(mode = "search",
                           term = checkmate::assert_string(term)),
              times = 5L) %>%
    # ensure we actually got a JSON response
    pal::assert_mime_type(mime_type = "application/json",
                          msg_suffix = mime_error_suffix) %>%
    # parse response
    httr::content(as = "parsed") %$%
    items %>%
    purrr::flatten_chr() 
}
```

# Referendum metadata

## `data_codebook`

```{r}
#' C2D Codebook
#'
#' A tibble containing the complete metadata of all [referendums()] variables. The Codebook below is also available
#' [online](https://rpkg.dev/c2d/articles/codebook.html).
#'
#' @includeRmd vignettes/codebook.Rmd
#'
#' @format `r pkgsnip::return_label("data")`
#' @aliases codebook
#' @family metadata
#' @export
"data_codebook"
```

## `val_lbls`

```{r}
#' Get the set of possible *value labels* of a referendum data variable
#'
#' Returns a character vector of value labels of a specific [referendums()] column, in the same order as [v_vals()], or of length `0` if `v_name`'s values are
#' not restricted to a predefined set or no value labels are defined in the [codebook][data_codebook].
#'
#' @param v_name A variable name. Must be one of the column names of [`data_codebook`].
#' @param incl_affixes Whether or not to add the corresponding `value_label_prefix` and `value_label_suffix` to the returned labels.
#'
#' @return A character vector. Of length `0` if `v_name`'s values are not restricted to a predefined set or no value labels are defined in the
#'   [codebook][data_codebook].
#' @seealso [`data_codebook`][data_codebook]
#' @family metadata
#' @export
#'
#' @examples
#' c2d::val_lbls("result",
#'               incl_affixes = FALSE)
#' c2d::val_lbls("result")
#' 
#' # Convert the labels to sentence case with trailing dot
#' c2d::val_lbls("result") |> pal::sentenceify()
val_lbls <- function(v_name,
                     incl_affixes = TRUE) {
  v_name <- rlang::arg_match(arg = v_name,
                             values = data_codebook$variable_name)
  metadata <- data_codebook %>% dplyr::filter(variable_name == !!v_name)
  result <- metadata$value_labels %>% purrr::flatten_chr()
  
  if (incl_affixes) {
    if (!is.na(metadata$value_label_prefix)) result <- paste(metadata$value_label_prefix, result)
    if (!is.na(metadata$value_label_suffix)) result <- paste(metadata$value_label_suffix, result)
  }
  
  result
}
```

## `v_vals`

```{r}
#' Get the set of possible *values* of a referendum data variable
#'
#' Returns a vector of the possible predefined values a specific column in [referendums()] can hold. If the variable values aren't restricted to a predefined
#' set, `NULL` is returned.
#'
#' @param v_name A variable name. Must be one of the column names of [`data_codebook`].
#'
#' @return
#' If `v_name`'s values are restricted to a predefined set and
#' - `v_name` is *not* of type list, a vector of the same type as `v_name`.
#' - `v_name` is of type list, a vector of the same type as the elements of `v_name`.
#'
#' Else `NULL`.
#' @seealso [`data_codebook`][data_codebook]
#' @family metadata
#' @export
#'
#' @examples
#' v_vals("result")
#' v_vals("id")
v_vals <- function(v_name) {
  
  v_name <- rlang::arg_match(arg = v_name,
                             values = data_codebook$variable_name)
  data_codebook %>%
    dplyr::filter(variable_name == !!v_name) %$%
    variable_values %>%
    unlist()
}
```

# [Tags](https://rpkg.dev/c2d/articles/codebook.html#tags)

## `hierarchize_tags`

```{r}
#' Hierarchize Tags
#'
#' Reconstructs the hierarchical relations between the three tag variables `tags_tier_1`, `tags_tier_2` and `tags_tier_3`. Can also be used to simply determine
#' the parent tag(s) of any tag.
#'
#' @param x The tags to hierarchize. Either a character vector of tags or a single-row data frame containing at least the columns `tags_tier_1`, `tags_tier_2`
#'   and `tags_tier_3`.
#'
#' @return `r pkgsnip::param_label("data")`
#' @family tags
#' @export
#'
#' @examples
#' c2d::hierarchize_tags("territorial questions")
#'
#' # hierarchize the tags of all Austrian referendums
#' c2d::referendums(country_code = "AT") |>
#'   dplyr::group_split(id) |>
#'   purrr::map(c2d::hierarchize_tags)
hierarchize_tags <- function(x) {
  
  test_char <- checkmate::test_character(x, any.missing = FALSE)
  
  if (!test_char) {
    
    tag_v_names <- paste0("tags_tier_", 1:3)
    test_df <- checkmate::test_data_frame(x,
                                          min.rows = 1L,
                                          max.rows = 1L)
    has_tag_vars <- all(tag_v_names %in% colnames(x))
    
    if (!test_df || !has_tag_vars) {
      cli::cli_abort(paste0("{.arg x} must be either a character vector of tags or a single-row data frame containing at least the columns ",
                            "{.field tags_tier_1}, {.field tags_tier_2} and {.field tags_tier_3}."))
    }
    
    x %<>%
      .[, tag_v_names] %>%
      purrr::pmap(~ as.character(c(..1, ..2, ..3))) %>%
      purrr::flatten_chr()
  }
  
  checkmate::assert_subset(x,
                           choices = tags(),
                           empty.ok = TRUE,
                           .var.name = "tags")
  
  tags_tier_1 <- x[x %in% tags(tiers = 1L)]
  tags_tier_2 <- x[x %in% tags(tiers = 2L)]
  tags_tier_3 <- x[x %in% tags(tiers = 3L)]
  inferred_tags_tier_1 <- infer_tags(tags = c(tags_tier_2, tags_tier_3),
                                     tier = 1L)
  inferred_tags_tier_2 <- infer_tags(tags = tags_tier_3,
                                     tier = 2L)
  non_parent_tags_tier_1 <- tags_tier_1 %>% setdiff(inferred_tags_tier_1)
  non_parent_tags_tier_2 <- tags_tier_2 %>% setdiff(inferred_tags_tier_2)
  
  # 0. initialize empty tibble
  result <- tibble::tibble(tag_tier_1 = character(),
                           tag_tier_2 = character(),
                           tag_tier_3 = character())
  
  # 1. add third-tier tags
  result <-
    tags_tier_3 %>%
    purrr::map_dfr(~ tibble::tibble(tag_tier_1 = infer_tags(tags = .x,
                                                            tier = 1L),
                                    tag_tier_2 = infer_tags(tags = .x,
                                                            tier = 2L),
                                    tag_tier_3 = .x)) %>%
    dplyr::bind_rows(result)
  
  # 2. add remaining second-tier tags
  if (length(non_parent_tags_tier_2)) {
    
    result <-
      non_parent_tags_tier_2 %>%
      purrr::map_dfr(~ tibble::tibble(tag_tier_1 = infer_tags(tags = .x,
                                                              tier = 1L),
                                      tag_tier_2 = .x,
                                      tag_tier_3 = NA_character_)) %>%
      dplyr::bind_rows(result)
  }
  
  # 3. add remaining top-tier tags
  if (length(non_parent_tags_tier_1)) {
    
    result %<>% dplyr::bind_rows(tibble::tibble(tag_tier_1 = non_parent_tags_tier_1,
                                                tag_tier_2 = NA_character_,
                                                tag_tier_3 = NA_character_))
  }
  
  # sort result
  result %>% dplyr::arrange(tag_tier_1, tag_tier_2, tag_tier_3)
}
```

## `infer_tags`

```{r}
#' Infer higher-tier tags
#'
#' Determines the top-tier (`tier = 1L`) or second-tier (`tier = 2L`) tags that are the parents or grandparents of `tags` in the
#' [hierarchy](https://rpkg.dev/c2d/articles/codebook.html#tags).
#'
#' @param tags The tags to determine the corresponding (grand)parent tags of. A factor or character vector.
#' @param tier The tier of the inferred tags. Should be >= than the tier of the `tags`. An integer scalar.
#'
#' @return A character vector.
#' @family tags
#' @export
#'
#' @examples
#' c2d::infer_tags(tags = c("EU", "animal protection"),
#'                 tier = 1L)
#' c2d::infer_tags(tags = c("EU", "animal protection"),
#'                 tier = 2L)
#' 
#' # tags of different tiers can mixed in `tags`
#' c2d::infer_tags(tags = c("EU", "environment"),
#'                 tier = 2L)
#' 
#' # but `tags` of a higher tier than `tier` will be ignored
#' c2d::infer_tags(tags = "foreign policy",
#'                 tier = 2L)
infer_tags <- function(tags,
                       tier = 1L) {
  
  if (is.factor(tags)) tags <- as.character(tags)
  
  checkmate::assert_subset(tags,
                           choices = tags())
  checkmate::assert_int(tier,
                        lower = 1L,
                        upper = 2L)
  
  # inferred from lower-tier tags
  result <- data_tags[data_tags$tag_tier_2 %in% tags | data_tags$tag_tier_3 %in% tags, ]
  result %<>% .[[paste0("tag_tier_", tier)]]
  
  # plus top-tier tags
  if (tier == 1L) result %<>% c(tags[tags %in% tags(tiers = 1L)])
  
  unique(result)
}
```

# C2D API Status

## `is_online`

```{r}
#' Test C2D API availability
#'
#' Checks if the C2D API server is online and operational.
#'
#' @param quiet Whether or not to suppress printing a warning in case the API is unavailable.
#'
#' @return A logical scalar.
#' @family api_status
#' @export
is_online <- function(quiet = FALSE) {
  
  checkmate::assert_flag(quiet)
  result <- FALSE
  response <-
    rlang::with_handlers(.expr = httr::RETRY(verb = "GET",
                                             url = "https://services.c2d.ch/health",
                                             times = 5L),
                         error = ~ paste0("Error: ", .x$message)) %>%
    # ensure we actually got a plaintext response
    pal::assert_mime_type(mime_type = "text/plain",
                          msg_suffix = mime_error_suffix) %>%
    # parse response
    httr::content(as = "text",
                  encoding = "UTF-8")
  
  if (response == "OK") {
    result <- TRUE
    
  } else if (!quiet) {
    cli::cli_warn("C2D API server responded with: {.val {response}}")
  }
  
  result
}
```
