---
editor_options:
  chunk_output_type: console
---

# NOTES

The C2D Database offers a [privately documented](https://github.com/ccmdesign/c2d-app/blob/master/docs/services.md) API available under `services.c2d.ch` with
the following "endpoints":

-   `referendums`: get referendum data

    Data on individual referendums can be retrieved under `referendums/ID` where `ID` stands for the referendum's unique identifier `oid` (which corresponds to
    the column `id` in the data returned by the `referendums()` function of this package).

    It supports the following URL parameters:

    -   `mode`: the operation mode; one of `stream`, `count` or `search`
    -   `term` the search term used for `mode=search`; note that only the `title$en` field will be searched
    -   `format`: the data format to be returned when `mode=stream`; one of `json` or `csv`
    -   `filter`: a valid [JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents) to customize queries in the
        underlying MongoDB collection; needs to be Base64-encoded; there are various [query
        operators](https://docs.mongodb.com/manual/reference/operator/query/) available to build complex queries

    The following values are used to denote empty values in the returned data:

    -   In integer fields:

        -   `-1L` (means "information not available or nonexistent")
        -   `-2L` ([means "information not given"](https://github.com/ccmdesign/c2d-app/issues/13#issuecomment-802466615))

    -   In character fields:

        -   `""`
        -   `"-1"`
        -   `"unknown"` (only for certain fields)
        -   `"-2"`
        -   `"not provided"` (only for certain fields)

    -   In array fields:

        -   (Named) `list()` in array fields

-   The returned JSON data has between 22 and 26 fields. Fields that are not always present include:

    -   `archive`
    -   `canton`
    -   `municipality`
    -   `number`

-   `referendums/stats`: get the number of referendums

    It accepts the following URL parameter:

    -   `filter`: a valid [JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents) to customize queries in the
        underlying MongoDB collection; needs to be Base64-encoded; there are various [query
        operators](https://docs.mongodb.com/manual/reference/operator/query/) available to build complex queries

-   `s3_objects`: retrieve a file stored on the C2D [Amazon S3 bucket](https://en.wikipedia.org/wiki/Amazon_S3#Design)

    It's not a real REST API endpoint but rather a URL redirection (I guess). It expects an object key to be appended like this:

    ``` {.http}
    GET https://services.c2d.ch/s3_objects/referendum_5f859fafd1291cc3961f1bc2_0002.pdf
    ```

-   `states`: get the first 10 sub-national states that match the query

    It accepts the following URL parameters to customize the query:

    -   `country`: limits the results to a specific country code
    -   `term`: limits the result to states matching this search term

-   `health`: check API status

# INTERNAL

## Avoid `R CMD check` notes about undefined global objects used in magrittr pipes

Cf. <https://github.com/tidyverse/magrittr/issues/29#issuecomment-74313262>

```{r}
utils::globalVariables(names = c(".",
                                 "all_of",
                                 "any_of",
                                 "applicability_constraint",
                                 "archive",
                                 "committee_name",
                                 "content-disposition",
                                 "country_code",
                                 "country_name",
                                 "date_time_created",
                                 "electorate_abroad",
                                 "electorate_total",
                                 "files",
                                 "id",
                                 "id_official",
                                 "id_sudd",
                                 "id_sudd_prefix",
                                 "is_draft",
                                 "items",
                                 "level",
                                 "municipality",
                                 "remarks",
                                 "result",
                                 "rowid",
                                 "subnational_entity_name",
                                 "sudd_prefix",
                                 "tag_tier_2",
                                 "tag_tier_3",
                                 "tags_tier_1",
                                 "tags_tier_2",
                                 "tags_tier_3",
                                 "title_de",
                                 "title_fr",
                                 "title_en",
                                 "type",
                                 "url_sudd",
                                 "url_swissvotes",
                                 "value_labels",
                                 "variable_name",
                                 "variable_values",
                                 "votes",
                                 "where"))
```

## Package load/unload

```{r}
.onLoad <- function(libname, pkgname) {
  pkgpins::clear(pkg = pkgname,
                 max_age = getOption("c2d.max_cache_lifespan",
                                     default = "30 days"))
}

.onUnload <- function(libpath) {
  pkgpins::deregister(pkg = pkg)
}
```

## Create custom CSS file for package vignettes

Assembles a custom CSS file by combining [rmarkdown's default lightweight vignette
stylesheet](https://github.com/rstudio/rmarkdown/blob/master/inst/rmarkdown/templates/html_vignette/resources/vignette.css) with our custom pkgdown CSS.

The R code chunk below is *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) (`purl = FALSE`) and thus has
to be manually executed in order to regenerate the custom CSS file.

```{r, purl = FALSE}
pal::assert_pkg("brio")

fs::path_package("rmarkdown/templates/html_vignette/resources/vignette.css",
                 package = "rmarkdown") %>%
  brio::read_file() %>%
  paste0("/* BEGIN custom CSS */\n",
         brio::read_file("pkgdown/extra.css"),
         "/* END custom CSS */\n") %>%
  brio::write_file(path = "vignettes/custom.css")
```

## Functions

### `httr_config`

Until [recently](https://github.com/ccmdesign/c2d-app/issues/10), the C2D API server didn't provide the [intermediate *R3*
certificate](https://crt.sh/?q=+730c1bdcd85f57ce5dc0bba733e5f1ba5a925b2a771d640a26f7a454224dad3b) (issued by Let's Encrypt) and since [curl doesn't support
*Authority Information Access* yet](https://github.com/curl/curl/issues/2793) to automatically discover it, we had to manually specify it. Thus, the certificate
was shipped with this package (found under `certs/3479778542.crt`). Because this isn't needed anymore, the cert as well as the code chunk below (`purl = FALSE`)
are ignored from package builds but kept in case they are needed again at some point in the future.

```{r, purl = FALSE}
httr_config <- function() {
  
  httr::config(cainfo = fs::path_package(package = pkg,
                                         "certs", "3479778542.crt"))
}
```

### `assemble_query_filter`

NOTE: `date` is not stored as an actual [MongoDB `Date` type](https://docs.mongodb.com/manual/reference/bson-types/#std-label-document-bson-type-date) in the
database, so we can't filter on it. It has [been requested](https://github.com/ccmdesign/c2d-app/issues/20) to change this upstream.

```{r}
#' Assemble MongoDB query filter document
#'
#' @param country_code The `country_code`(s) to be included. A character vector.
#' @param subnational_entity_name The `subnational_entity_name`(s) to be included. A character vector.
#' @param municipality The `municipality`(s) to be included. A character vector.
#' @param level The `level`(s) to be included. A character vector.
#' @param type The `type`(s) to be included. A character vector.
#' @param date The `date`(s) to be included. A character vector.
#' @param is_draft `TRUE` means to include only referendum entries with _draft_ status, `FALSE` to include only normal entries. Set to `NULL` in order to
#'   include both draft and normal entries.
#' @param date_time_created_min The minimum `date_time_created` to be included. A [date][base::Date], a [datetime][base::DateTimeClasses], or something
#'   coercible to (like `"2006-01-02"` or `"2006-01-02T15:04:05"`).
#' @param date_time_created_max The maximum `date_time_created` to be included. A [date][base::Date], a [datetime][base::DateTimeClasses], or something
#'   coercible to (like `"2006-01-02"` or `"2006-01-02T15:04:05"`).
#' @param query_filter A valid [MongoDB JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents) which allows for
#'   maximum control over what data is included. This takes precedence over all the other parameters, i.e. if `query_filter` is provided, the parameters
#'   ``r formals(assemble_query_filter) %>% names() %>% setdiff(c("query_filter", "base64_encode")) %>% pal::prose_ls(wrap = "`")`` are ignored.
#' @param base64_encode Whether or not to [Base64](https://en.wikipedia.org/wiki/Base64)-encode the resulting query filter document. Note that the
#'   `query_filter` argument provided to other functions of this package must be Base64-encoded.
#'
#' @return A character scalar containing a valid [MongoDB JSON query filter document](https://docs.mongodb.com/manual/core/document/#query-filter-documents),
#'   [Base64](https://en.wikipedia.org/wiki/Base64)-encoded if `base64_encode = TRUE`.
#' @keywords internal
assemble_query_filter <- function(country_code = NULL,
                                  subnational_entity_name = NULL,
                                  municipality = NULL,
                                  level = NULL,
                                  type = NULL,
                                  date = NULL,
                                  is_draft = NULL,
                                  date_time_created_min = NULL,
                                  date_time_created_max = NULL,
                                  query_filter = NULL,
                                  base64_encode = TRUE) {
  
  # assemble JSON query filter document if `query_filter` is not provided
  if (is.null(checkmate::assert_string(query_filter, null.ok = TRUE))) {
    
    query_filter <-
      list(country_code =
             country_code %>%
             purrr::map_chr(checkmate::assert_choice,
                            choices = countrycode::codelist$iso2c %>% setdiff(NA_character_),
                            null.ok = TRUE,
                            .var.name = "country_code") %>%
             query_filter_in(),
           canton = query_filter_in(checkmate::assert_character(subnational_entity_name,
                                                                any.missing = FALSE,
                                                                null.ok = TRUE)),
           municipality = query_filter_in(checkmate::assert_character(municipality,
                                                                      any.missing = FALSE,
                                                                      null.ok = TRUE)),
           level =
             level %>%
             purrr::map_chr(checkmate::assert_choice,
                            choices = v_vals("level"),
                            null.ok = TRUE,
                            .var.name = "level") %>%
             purrr::when(length(.) == 0L ~ .,
                         ~ dplyr::recode(., "subnational" = "sub-national") %>% stringr::str_to_sentence()) %>%
             query_filter_in(),
           institution =
             type %>%
             purrr::map_chr(checkmate::assert_choice,
                            choices = v_vals("type"),
                            null.ok = TRUE,
                            .var.name = "type") %>%
             purrr::when(length(.) == 0L ~ .,
                         ~ dplyr::recode("citizens' assembly" = "citizen assembly") %>% stringr::str_to_sentence()) %>%
             query_filter_in(),
           date = query_filter_in(checkmate::assert_character(as.character(date),
                                                              pattern = "\\d{4}-\\d{2}-\\d{2}",
                                                              any.missing = FALSE,
                                                              null.ok = TRUE,
                                                              .var.name = "date")),
           draft = checkmate::assert_flag(is_draft,
                                          null.ok = TRUE),
           created_on = query_filter_date(min = checkmate::assert_posixct(lubridate::as_datetime(date_time_created_min),
                                                                          any.missing = FALSE,
                                                                          .var.name = "date_time_created_min"),
                                          max = checkmate::assert_posixct(lubridate::as_datetime(date_time_created_max),
                                                                          any.missing = FALSE,
                                                                          .var.name = "date_time_created_max"))) %>%
      # remove `NULL` elements
      purrr::compact() %>%
      # convert to base64-encoded JSON
      jsonlite::toJSON(POSIXt = "ISO8601",
                       auto_unbox = TRUE,
                       digits = NA,
                       pretty = FALSE)
  }
  
  query_filter %>% purrr::when(checkmate::assert_flag(base64_encode) ~ jsonlite::base64_enc(.),
                               ~ .)
}
```

### `read_toml`

```{r}
#' Read in a TOML file
#'
#' @param path Path to the TOML file. A [path][fs::fs_path] or something coercible to.
#'
#' @return The TOML file's content as a [strict list][xfun::strict_list].
#' @keywords internal
read_toml <- function(path) {
  
  pal::assert_pkg("RcppTOML")
  pal::assert_pkg("xfun")
  
  path %>%
    purrr::when(length(.) > 0L ~ RcppTOML::parseTOML(input = .,
                                                     escape = FALSE),
                ~ NULL) %>%
    xfun::as_strict_list()
}
```

### `tidy_referendums`

```{r}
#' Tidy "raw" MongoDB referendum data
#'
#' Converts from the "raw" MongoDB to the tidied `referendums()` schema.
#'
#' @param data The data from the C2D MongoDB as a list (converted from the JSON).
#' @param tidy Whether or not to tidy the referendum data, i.e. apply various data cleansing tasks and add additional variables. If `FALSE`, the raw MongoDB
#'   referendum data will only be modified just enough to be able to return it as a tibble. Note that the latter doesn't conform to the [codebook][codebook]
#'   (i.a. different variable names).
#'
#' @return `r pkgsnip::return_label("data")`
#' @keywords internal
tidy_referendums <- function(data,
                             tidy = TRUE) {
  data %<>%
    # unnest columns and ensure list type for multi-value columns
    purrr::map(.f = function(l,
                             category_names = names(l$categories),
                             context_names = names(l$context),
                             title_langs = names(l$title)) {
      
      for (name in category_names) {
        l[[name]] <- l$categories[[name]]
      }
      
      for (name in context_names) {
        l[[name]] <- l$context[[name]]
      }
      
      for (lang in title_langs) {
        l[[paste0("title_", lang)]] <- l$title[[lang]]
      }
      
      l$categories <- NULL
      l$context <- NULL
      l$title <- NULL
      l$`_id` %<>% purrr::flatten_chr()
      
      for (name in c("action",
                     "tags",
                     "special_topics",
                     "excluded_topics")) {
        l[[name]] %<>% purrr::flatten_chr() %>% list()
      }
      
      for (name in c("archive",
                     "files",
                     "votes_per_canton")) {
        l[[name]] %<>% list()
      }
      
      l
    }) %>%
    # convert to tibble
    purrr::map_dfr(tibble::as_tibble_row)
  
  # check data integrity
  ## ensure mandatory fields are never `NA`
  if (data %>%
      dplyr::select(all_of(required_fields)) %>%
      dplyr::filter(dplyr::if_any(.fns = is.na)) %>%
      nrow()) {
    
    cli::cli_abort("Mandatory fields detected that contain {.val NA} values. Please debug.")
  }
  
  # ## ensure `oid`s are unique
  # if (anyDuplicated(data$`_id`)) {
  #   cli::cli_abort("Duplicated {.var oid}s detected. Please debug.")
  # }
  
  ## ensure `country_codes`s are valid
  invalid_country_codes <- data$country_code[!(data$country_code %in% countrycode::codelist$iso2c %>% setdiff(NA_character_))]
  
  if (length(invalid_country_codes)) {
    cli::cli_abort("Unknown {.var country_code}s detected. Please debug.")
  }
  
  # tidy data
  if (tidy) {
    
    data %<>%
      # rename variables (mind that the MongoDB-based API doesn't demand a schema)
      dplyr::rename(!!!({
        nms <- names(.)
        v_names %>% purrr::keep(~ .x %in% nms)
      })) %>%
      # add vars which aren't always included
      dplyr::full_join(y = tibble::tibble(archive = list(),
                                          subnational_entity_name = character(),
                                          municipality = character(),
                                          id_official = character()),
                       by = intersect(c("archive",
                                        "subnational_entity_name",
                                        "municipality",
                                        "id_official"),
                                      colnames(.))) %>%
      
      # create/recode variables
      dplyr::mutate(
        # replace empty lists with `NULL` in list columns
        dplyr::across(where(is.list),
                      ~ dplyr::case_when(.x %>% purrr::map_lgl(~ length(.x) == 0L) ~ list(NULL),
                                         TRUE ~ .x)),
        
        # ensure all supposed to floating-point numbers are actually of type double (JSON API is not reliable in this respect)
        dplyr::across(any_of(c("subterritories_no",
                               "subterritories_yes",
                               "date_time_created")),
                      as.double),
        
        # use explicit NA values
        dplyr::across(where(is.integer),
                      ~ dplyr::if_else(.x %in% c(-1L, -2L),
                                       NA_integer_,
                                       .x)),
        dplyr::across(where(is.character),
                      ~ dplyr::if_else(.x %in% c("", "-1", "-2"),
                                       NA_character_,
                                       .x)),
        dplyr::across(any_of(c("subterritories_yes", "subterritories_no")),
                      ~ dplyr::if_else(.x %in% c(-1.0, -2.0),
                                       NA_real_,
                                       .x)),
        result = result %>% dplyr::if_else(. %in% c("Unknown", "Not provided"),
                                           NA_character_,
                                           .),
        
        # convert all values to lowercase
        ## vectors
        dplyr::across(c(level,
                        result,
                        type,
                        any_of(c("inst_legal_basis_type",
                                 "inst_object_type",
                                 "inst_object_legal_level",
                                 "inst_object_revision_extent",
                                 "inst_trigger_type",
                                 "inst_trigger_actor_level",
                                 "inst_trigger_time_limit",
                                 "inst_quorum_approval",
                                 "inst_precondition_actor",
                                 "inst_precondition_decision"))),
                      stringr::str_to_lower),
        ## lists
        dplyr::across(.cols = any_of(c("inst_topics_only",
                                       "inst_topics_excluded")),
                      .fns = purrr::map,
                      .f = stringr::str_to_lower),
        
        # only convert non-abbreviations to lowercase
        ## vectors
        dplyr::across(.cols = any_of(c("inst_object_author",
                                       "inst_trigger_actor",
                                       "inst_precondition_actor")),
                      .fns = purrr::map_chr,
                      .f = lower_non_abbrs),
        ## lists
        dplyr::across(.cols = tags,
                      .fns = purrr::map,
                      .f = clean_tags),
        
        # specific recodings
        ## binary (dummies)
        dplyr::across(any_of("position_government"),
                      ~ dplyr::case_when(.x == "Acceptance" ~ "yes",
                                         .x == "Rejection" ~ "no",
                                         TRUE ~ NA_character_)),
        dplyr::across(any_of("inst_has_urgent_legal_basis"),
                      ~ dplyr::case_when(.x == "Urgent" ~ TRUE,
                                         .x == "Normal" ~ FALSE,
                                         TRUE ~ NA)),
        dplyr::across(any_of("inst_is_binding"),
                      ~ dplyr::case_when(.x == "Binding" ~ TRUE,
                                         .x == "Non-binding" ~ FALSE,
                                         TRUE ~ NA)),
        dplyr::across(any_of("inst_is_counter_proposal"),
                      ~ dplyr::case_when(.x == "Yes" ~ TRUE,
                                         .x == "No" ~ FALSE,
                                         TRUE ~ NA)),
        dplyr::across(any_of("inst_is_assembly"),
                      ~ dplyr::case_when(.x == "Assembly" ~ TRUE,
                                         .x == "Ballot" ~ FALSE,
                                         TRUE ~ NA)),
        dplyr::across(any_of("inst_has_precondition"),
                      ~ dplyr::case_when(.x == "Exists" ~ TRUE,
                                         .x == "Does not exist" ~ FALSE,
                                         TRUE ~ NA)),
        ## nominal
        ### split `id_official` into `id_sudd` (the latter consists of a two-letter country code plus a 6-digit number)
        id_sudd =
          id_official %>%
          dplyr::if_else(stringr::str_detect(string = .,
                                             pattern = "^\\D"),
                         .,
                         NA_character_) %>%
          # everything beyond the 8th char seems to be manually added -> strip!
          stringr::str_sub(end = 8L),
        id_official =
          id_official %>%
          dplyr::if_else(stringr::str_detect(string = .,
                                             pattern = "^\\D"),
                         NA_character_,
                         .),
        ### split `tags` into separate per-tier vars
        tags_tier_1 = tags %>% purrr::map(infer_tags,
                                          tier = 1L),
        tags_tier_2 = tags %>% purrr::map(infer_tags,
                                          tier = 2L),
        tags_tier_3 = tags %>% purrr::map(~ .x[.x %in% tags_tier_3_cleaned]),
        ### various cleanups
        level = stringr::str_replace(string = level,
                                     pattern = "sub-national",
                                     replacement = "subnational"),
        type = type %>% dplyr::recode("citizen assembly" = "citizens' assembly",
                                      "not provided" = NA_character_),
        dplyr::across(any_of("inst_trigger_actor"),
                      dplyr::recode,
                      "institution" = "other institution"),
        dplyr::across(any_of("inst_object_type"),
                      dplyr::recode,
                      "legal text (ausformulierter vorschlag)" = "legal text (formulated proposal)",
                      "legal text (allg. anregung)" = "legal text (general proposal)"),
        dplyr::across(any_of("inst_object_revision_divisibility"),
                      dplyr::recode,
                      "Whole text only" = "as a whole",
                      "Splitting up possible" = "split up",
                      "Variants possible" = "as variants",
                      "Variants / splitting up possible" = "both split up and as variants"),
        dplyr::across(.cols = any_of("inst_topics_only"),
                      .fns = purrr::map,
                      .f = dplyr::recode,
                      "infrastructural act" = "infrastructural acts",
                      "competence shift" = "competence shifts",
                      "financial act" = "financial acts",
                      "financial act (expenses)" = "financial acts (expenses)",
                      "financial act (taxes)" = "financial acts (taxes)",
                      "financial act (obligations)" = "financial acts (obligations)",
                      "total revision of the constitution" = "total revisions of the constitution"),
        dplyr::across(.cols = any_of("inst_topics_excluded"),
                      .fns = purrr::map,
                      .f = dplyr::recode,
                      "budget" = "budgets"),
        dplyr::across(any_of("inst_quorum_turnout"),
                      stringr::str_replace_all,
                      pattern = c("^(\\s+)?>(\\s+)?" = ">\u202f",
                                  "(\\s+)?%(\\s+)?$" = "\u202f%")),
        ## ordinal
        ## interval
        date = lubridate::as_date(date),
        date_time_created = date_time_created %>% magrittr::divide_by(1000L) %>% lubridate::as_datetime(),
        ## undefined
        files = files %>% purrr::map(~ .x %>%
                                       # harmonize depth (i.e. add list level if necessary)
                                       purrr::when(length(.) == 0L || purrr::vec_depth(.) > 3L ~ .,
                                                   ~ list(.)) %>%
                                       purrr::map(~ .x %>%
                                                    # unnest and restore `date`
                                                    purrr::assign_in(where = "date",
                                                                     value =
                                                                       .$date %>%
                                                                       purrr::flatten_dbl() %>%
                                                                       magrittr::divide_by(1000L) %>%
                                                                       lubridate::as_datetime()) %>%
                                                    # change subvariable names
                                                    magrittr::set_names(names(.) %>% dplyr::recode("date" = "date_time_attached",
                                                                                                   "object_key" = "s3_object_key",
                                                                                                   "size" = "file_size",
                                                                                                   "deleted" = "is_deleted")))),
        
        # variable creations
        url_sudd = dplyr::if_else(!is.na(id_sudd),
                                  paste0("https://sudd.ch/event.php?id=", id_sudd),
                                  NA_character_),
        url_swissvotes = dplyr::if_else(country_code == "CH" & level == "national" & !is.na(id_official),
                                        paste0("https://swissvotes.ch/vote/", id_official),
                                        NA_character_)
      ) %>%
      
      # remove obsolete vars
      dplyr::select(-tags) %>%
      
      # convert types
      ## based on codebook
      dplyr::mutate(dplyr::across(.fns = ~ {
        
        metadata <- data_codebook %>% dplyr::filter(variable_name == dplyr::cur_column())
        if (nrow(metadata) != 1L) cli::cli_abort("Missing codebook metadata! Please debug")
        
        if (!is.logical(.x) && !is.null(metadata$variable_values[[1L]])) {
          
          lvls <- purrr::flatten_chr(metadata$variable_values)
          is_ordered <- metadata$value_scale %in% c("ordinal_ascending", "ordinal_descending")
          
          if (is.list(.x)) {
            .x %>% purrr::map(.f = factor,
                              levels = lvls,
                              ordered = is_ordered)
          } else {
            factor(x = .x,
                   levels = lvls,
                   ordered = is_ordered)
          }
        } else .x
      })) %>%
      ## nominal vars without variable_values in codebook
      dplyr::mutate(dplyr::across(any_of(c("country_code",
                                           "country_name",
                                           "subnational_entity_name",
                                           "municipality")),
                                  as.factor)) %>%
      
      # add variable labels
      labelled::set_variable_labels(.labels = var_lbls,
                                    .strict = FALSE) %>%
      
      # reorder columns
      dplyr::relocate(id,
                      id_official,
                      id_sudd,
                      country_code,
                      country_name,
                      subnational_entity_name,
                      municipality,
                      level,
                      date,
                      title_de,
                      title_fr,
                      title_en,
                      committee_name,
                      result,
                      any_of("subterritories_yes"),
                      any_of("subterritories_no"),
                      electorate_total,
                      electorate_abroad,
                      any_of("votes_yes"),
                      any_of("votes_no"),
                      any_of("votes_empty"),
                      any_of("votes_invalid"),
                      any_of("votes_per_subterritory"),
                      any_of("lower_house_yes"),
                      any_of("lower_house_no"),
                      any_of("lower_house_abstentions"),
                      any_of("upper_house_yes"),
                      any_of("upper_house_no"),
                      any_of("upper_house_abstentions"),
                      any_of("position_government"),
                      tags_tier_1,
                      tags_tier_2,
                      tags_tier_3,
                      remarks,
                      files,
                      url_sudd,
                      url_swissvotes,
                      is_draft,
                      date_time_created,
                      archive,
                      type,
                      any_of("inst_legal_basis_type"),
                      any_of("inst_has_urgent_legal_basis"),
                      any_of("inst_is_binding"),
                      any_of("inst_is_counter_proposal"),
                      any_of("inst_is_assembly"),
                      any_of("inst_trigger_type"),
                      any_of("inst_trigger_actor"),
                      any_of("inst_trigger_actor_level"),
                      any_of("inst_trigger_threshold"),
                      any_of("inst_trigger_time_limit"),
                      any_of("inst_object_type"),
                      any_of("inst_object_author"),
                      any_of("inst_object_legal_level"),
                      any_of("inst_object_revision_extent"),
                      any_of("inst_object_revision_modes"),
                      any_of("inst_object_revision_divisibility"),
                      any_of("inst_topics_only"),
                      any_of("inst_topics_excluded"),
                      any_of("inst_quorum_turnout"),
                      any_of("inst_quorum_approval"),
                      any_of("inst_has_precondition"),
                      any_of("inst_precondition_actor"),
                      any_of("inst_precondition_decision"))
    
    # correct known errors
    ## `id_sudd`
    data$id_sudd[data$country_code == "AZ" & data$id_sudd == "az021009"] <- "az022009"
    data$id_sudd[data$country_code == "BS" & data$id_sudd == "bs021002"] <- "bs022002"
    data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011972"] <- "mg011972"
    data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011975"] <- "mg011975"
    data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011992"] <- "mg011992"
    data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011995"] <- "mg011995"
    data$id_sudd[data$country_code == "MG" & data$id_sudd == "md011998"] <- "mg011998"
    data$id_sudd[data$country_code == "MG" & data$id_sudd == "md012007"] <- "mg012007"
    data$id_sudd[data$country_code == "MG" & data$id_sudd == "md012010"] <- "mg012010"
    data$id_sudd[data$country_code == "MH" & data$id_sudd == "mh00578"] <- "mh011979"
    data$id_sudd[data$country_code == "MW" & data$id_sudd == "mw012003"] <- "mw011993"
    data$id_sudd[data$country_code == "EC" & data$id_sudd == "ec031993"] <- "ec031994"
    data$id_sudd[data$country_code == "EC" & data$id_sudd == "ec021006"] <- "ec022006"
    data$id_sudd[data$country_code == "GQ" & data$id_sudd == "gq012968"] <- "gq011968"
    data$id_sudd[data$country_code == "GT" & data$id_sudd == "gt011998"] <- "gt011999"
    data$id_sudd[data$country_code == "GT" & data$id_sudd == "gt021998"] <- "gt021999"
    data$id_sudd[data$country_code == "GT" & data$id_sudd == "gt031998"] <- "gt031999"
    data$id_sudd[data$country_code == "LI" & data$id_sudd == "li021920"] <- "li021930"
    data$id_sudd[data$country_code == "LT" & data$id_sudd == "lt031993"] <- "lt031994"
    data$id_sudd[data$country_code == "NF" & data$id_sudd == "nf011980"] <- "nf011979"
    data$id_sudd[data$country_code == "TV" & data$id_sudd == "tv011994"] <- "tv011974"
    data$id_sudd[data$country_code == "UY" & data$id_sudd == "uy011918"] <- "uy011917"
    data$id_sudd[data$country_code == "YE" & data$id_sudd == "ye012991"] <- "ye011991"
    # data$id_sudd[data$country_code == "" & data$id_sudd == ""] <- ""
  }
  
  data
}
```

### MongoDB query filter operator helpers

#### `$date`

Example query found in the mongolite documentation: <https://jeroen.github.io/mongolite/query-data.html#select-by-date>

```{r}
query_filter_date <- function(min,
                              max) {
  
  list(`$gte` = list(`$date` = lubridate::as_datetime(min)) %>% purrr::compact(),
       `$lte` = list(`$date` = lubridate::as_datetime(max)) %>% purrr::compact()) %>%
    purrr::compact()
}
```

#### `$in`

MongoDB documentation: <https://docs.mongodb.com/manual/reference/operator/query/in/#mongodb-query-op.-in>

```{r}
query_filter_in <- function(x) {
  
  x %>% purrr::when(is.null(.) || length(.) == 0L ~ NULL,
                    length(.) == 1 ~ .,
                    ~ list(`$in` = .))
}
```

### `lower_non_abbrs`

```{r}
#' Lowercase non-abbreviations
#'
#' @param x A character vector.
#'
#' @return A character vector of the same lenght as `x`.
#' @keywords internal
lower_non_abbrs <- function(x) {
  
  x %>%
    stringr::str_split(pattern = "\\s") %>%
    purrr::map_chr(~ .x %>%
                     dplyr::if_else(stringr::str_detect(string = .,
                                                        pattern = "^[^[:lower:]]+$"),
                                    .,
                                    stringr::str_to_lower(.)) %>%
                     paste0(collapse = " "))
}
```

### Tags-related

#### `clean_tags`

TODO:

-   PR to implement this in the back-end/MongoDB [was submitted](https://github.com/ccmdesign/c2d-app/pull/40); once merged, remove from this pkg!

```{r}
clean_tags <- function(tags) {
  
  lower_non_abbrs(tags) %>% dplyr::recode("armed forces - general" = "armed forces in general",
                                          "citizen's initiatives" = "citizens' initiatives",
                                          "european policy" = "European policy",
                                          "other european organisations" = "other European organisations",
                                          "religion - churches" = "religion, churches",
                                          "swiss abroad" = "Swiss abroad",
                                          "territoral questions" = "territorial questions")
}
```

#### `tags`

```{r}
tags <- function(tiers = 1:3) {
  
  checkmate::assert_integerish(tiers,
                               lower = 1L,
                               upper = 3L,
                               any.missing = FALSE,
                               unique = TRUE)
  tag_set <- character()
  
  if (1L %in% tiers) {
    tag_set %<>% c(data_tags_tidy$tag_tier_1)
  }
  
  if (2L %in% tiers) {
    tag_set %<>% c(data_tags_tidy$tag_tier_2)
  }
  
  if (3L %in% tiers) {
    tag_set %<>% c(data_tags_tidy$tag_tier_3)
  }
  
  tag_set %>% setdiff(NA_character_) %>% unique()
}
```

## Constants

```{r}
pkg <- utils::packageName()
```

### Mandatory data fields

Field names as returned by the `referendums` API endpoint.

```{r}
required_fields <- c("_id",
                     "country_name",
                     "created_on",
                     "level",
                     # "title_en", # contains NAs from before relaunch
                     # "type", # contains NAs from before relaunch
                     "total_electorate",
                     "votes_no",
                     "votes_yes")
```

### Variable renamings

This maps the final variable names as found in the [C2D codebook](https://rpkg.dev/c2d/articles/codebook.html) to the field names of the back-ends MongoDB. The
renamings should eventually all be upstreamed and this mapping shall be become superfluous.

```{r}
v_names <- list("id"                                = "_id",
                "id_official"                       = "number",
                "subnational_entity_name"           = "canton",
                "subterritories_no"                 = "states_no",
                "subterritories_yes"                = "states_yes",
                "electorate_total"                  = "total_electorate",
                "electorate_abroad"                 = "citizens_abroad",
                "votes_per_subterritory"            = "votes_per_canton",
                "lower_house_yes"                   = "national_council_yes",
                "lower_house_no"                    = "national_council_no",
                "lower_house_abstentions"           = "national_council_abstentions",
                "upper_house_yes"                   = "states_council_yes",
                "upper_house_no"                    = "states_council_no",
                "upper_house_abstentions"           = "states_council_abstentions",
                "position_government"               = "recommendation",
                "is_draft"                          = "draft",
                "date_time_created"                 = "created_on",
                "type"                              = "institution",
                "inst_legal_basis_type"             = "official_status",
                "inst_has_urgent_legal_basis"       = "legal_act_type",
                "inst_is_binding"                   = "vote_result_status",
                "inst_is_counter_proposal"          = "counter_proposal",
                "inst_is_assembly"                  = "vote_venue",
                "inst_trigger_type"                 = "vote_trigger",
                "inst_trigger_actor"                = "vote_trigger_actor",
                "inst_trigger_actor_level"          = "vote_trigger_state_level",
                "inst_trigger_threshold"            = "vote_trigger_number",
                "inst_trigger_time_limit"           = "vote_trigger_time",
                "inst_object_type"                  = "vote_object",
                "inst_object_author"                = "author_of_the_vote_object",
                "inst_object_legal_level"           = "hierarchy_of_the_legal_norm",
                "inst_object_revision_extent"       = "degree_of_revision",
                "inst_object_revision_modes"        = "action",
                "inst_object_revision_divisibility" = "referendum_text_options",
                "inst_quorum_turnout"               = "turnout_quorum",
                "inst_quorum_approval"              = "decision_quorum",
                "inst_has_precondition"             = "institutional_precondition",
                "inst_precondition_actor"           = "institutional_precondition_decision_actor",
                "inst_precondition_decision"        = "institutional_precondition_decision",
                "inst_topics_only"                  = "special_topics",
                "inst_topics_excluded"              = "excluded_topics")
```

## Data

The R code chunks below are *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) (`purl = FALSE`). The chunks
have to be executed at least once to generate the necessary internal data file [`R/sysdata.rda`](https://r-pkgs.org/data.html#data-sysdata).

### Codebook data

TODO: Once the toml\* package is finished, use it instead to do the stuff below more generically.

```{r, purl = FALSE}
data_codebook <- {
  
  pal::assert_pkg("assertr")
  codebook <- read_toml("data-raw/c2d_codebook.toml")
  
  c(codebook$`01_main`$item,
    codebook$`02_institutional`$`01_status`$item,
    codebook$`02_institutional`$`02_trigger`$item,
    codebook$`02_institutional`$`03_object`$item,
    codebook$`02_institutional`$`04_other`$item) %>%
    purrr::map_dfr(~ tibble::tibble(variable_name = .x$variable_name %||% NA_character_,
                                    variable_label = .x$variable_label %||% NA_character_,
                                    is_multi_valued = .x$is_multi_valued %||% FALSE,
                                    applicability_constraint = .x$applicability_constraint %||% NA_character_,
                                    variable_values = list(.x$variable_values) %||% list(character()),
                                    value_label_prefix = .x$value_label_prefix %||% NA_character_,
                                    value_labels = list(.x$value_labels) %||% list(character()),
                                    value_label_suffix = .x$value_label_suffix %||% NA_character_,
                                    value_default = list(.x$value_default) %||% list(character()),
                                    value_scale = .x$value_scale %||% NA_character_)) %>%
    # integrity checks
    assertr::assert(predicate = assertr::not_na,
                    variable_name,
                    variable_label,
                    value_scale,
                    is_multi_valued)
}
```

### Tags

```{r, purl = FALSE}
# download file from private CCM Design repo
data_tags <-
  pal::gh_text_file(owner = "ccmdesign",
                    name = "c2d-app",
                    rev = "master",
                    path = "ch.c2d/web/themes.json") %>%
  jsonlite::fromJSON(flatten = FALSE) %>%
  tibble::as_tibble()

# create tidy data version
data_tags_tidy <-
  data_tags %>%
  dplyr::rename(tag_tier_1 = name,
                tag_tier_2 = children) %>%
  tidyr::unnest_longer(col = tag_tier_2)

data_tags_tidy <-
  dplyr::bind_cols(tag_tier_1 = data_tags_tidy$tag_tier_1,
                   data_tags_tidy$tag_tier_2) %>%
  dplyr::rename(tag_tier_2 = name,
                tag_tier_3 = children) %>%
  # ensure consistent col subtypes
  dplyr::mutate(tag_tier_3 = tag_tier_3 %>% purrr::map(~ {if(is.list(.x)) character() else .x})) %>%
  tidyr::unnest_longer(col = tag_tier_3)
```

### Constants pre-computed for performance purposes

```{r, purl = FALSE}
data_tags_tidy_cleaned <- data_tags_tidy %>% dplyr::mutate(dplyr::across(.fns = clean_tags))
tags_cleaned <- clean_tags(tags())
tags_tier_1_cleaned <- clean_tags(tags(tiers = 1L))
tags_tier_2_cleaned <- clean_tags(tags(tiers = 2L))
tags_tier_3_cleaned <- clean_tags(tags(tiers = 3L))
# NOTE: `var_lbls` is assigned here because when inlined, it somehow breaks `devtools::run_examples()` during pkg check ("C stack usage (...) is too close to
#       the limit"); maybe too large fn stack? (modifying `pal::strip_md()` to minimize the fn stack was tried but didn't help)
var_lbls <-
  data_codebook$variable_label %>%
  pal::strip_md() %>%
  as.list() %>%
  magrittr::set_names(value = data_codebook$variable_name)
```

Finally, we have to save all the external data objects to `data/*.RData` as well as the internal data objects to `R/sysdata.rda`.

Note that `data_codebook` is saved as internal data and later exported because external data doesn't support `@aliases`.

```{r, purl = FALSE}
usethis::use_data(data_codebook,
                  data_tags,
                  data_tags_tidy,
                  data_tags_tidy_cleaned,
                  tags_cleaned,
                  tags_tier_1_cleaned,
                  tags_tier_2_cleaned,
                  tags_tier_3_cleaned,
                  var_lbls,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```

# `is_online`

```{r}
#' Test C2D API availability
#'
#' Checks if the C2D API server is online and operational.
#'
#' @param quiet Whether or not to suppress printing a warning in case the API is unavailable.
#'
#' @return A logical scalar.
#' @export
is_online <- function(quiet = FALSE) {
  
  checkmate::assert_flag(quiet)
  result <- FALSE
  response <- rlang::with_handlers(.expr = httr::RETRY(verb = "GET",
                                                       url = "https://services.c2d.ch/health",
                                                       times = 5L),
                                   error = ~ paste0("Error: ", .x$message))
  
  if (class(response) == "response") {
    
    response %<>% httr::content(as = "parsed",
                                encoding = "UTF-8")
    
    if (response == "OK") {
      result <- TRUE
      
    } else if (quiet) {
      cli::cli_warn("C2D API server responded with: {.val {response}}")
    }
  }
  
  result
}
```

# Referendums

## `referendums`

TODO:

-   [ ] Upstream renamings and recodings as far as possible and subsequently remove them from `referendums()` and `assemble_query_filter()`.

```{r}
#' Get referendum data
#'
#' Downloads the referendum data from the C2D Database. See [`codebook`][codebook] for a detailed description of all variables.
#'
#' @param use_cache `r pkgsnip::param_label("use_cache")`
#' @param cache_lifespan `r pkgsnip::param_label("cache_lifespan")`
#' @param incl_archive Whether or not to include an `archive` column containing data from an earlier, obsolete state of the C2D database.
#' @inheritParams assemble_query_filter
#' @inheritParams tidy_referendums
#'
#' @return `r pkgsnip::return_label("data")`
#' @export
#'
#' @examples
#' \dontrun{
#' # get all referendums (excl. drafts)
#' c2d::referendums()}
#' 
#' # get only referendums in Austria and Australia on subnational level
#' c2d::referendums(country_code = c("AT", "AU"),
#'                  level = "subnational")
#' 
#' # provide custom `query_filter` for more complex queries like regex matches
#' # cf. https://docs.mongodb.com/manual/reference/operator/query/regex/
#' c2d::referendums(query_filter = '{"country_code":{"$regex":"A."}}')
referendums <- function(use_cache = TRUE,
                        cache_lifespan = "1 week",
                        incl_archive = FALSE,
                        is_draft = FALSE,
                        country_code = NULL,
                        subnational_entity_name = NULL,
                        municipality = NULL,
                        level = NULL,
                        type = NULL,
                        date = NULL,
                        date_time_created_min = NULL,
                        date_time_created_max = NULL,
                        query_filter = NULL,
                        tidy = TRUE) {
  
  pkgpins::with_cache(.fn = ~ {
    
    checkmate::assert_flag(incl_archive)
    checkmate::assert_flag(tidy)
    
    # retrieve data
    data <-
      httr::RETRY(verb = "GET",
                  url = "https://services.c2d.ch/referendums",
                  query = list(mode = "stream",
                               format = "json",
                               filter = assemble_query_filter(country_code = country_code,
                                                              subnational_entity_name = subnational_entity_name,
                                                              municipality = municipality,
                                                              level = level,
                                                              type = type,
                                                              date = date,
                                                              is_draft = is_draft,
                                                              date_time_created_min = date_time_created_min,
                                                              date_time_created_max = date_time_created_max,
                                                              query_filter = query_filter)),
                  times = 5L) %>%
      httr::content(as = "parsed") %$%
      items %>%
      # tidy data
      tidy_referendums(tidy = tidy)
    
    # exclude `archive` if requested
    if (!incl_archive) data %<>% dplyr::select(-any_of("archive"))
    
    # return data
    data
  },
  .use_cache = use_cache,
  .cache_lifespan = cache_lifespan,
  .pkg = pkg)
}
```

## Metadata

### `data_codebook`

```{r}
#' C2D Codebook
#'
#' A tibble containing the complete metadata of all [referendums()] variables. The Codebook below is also available
#' [online](https://rpkg.dev/c2d/articles/codebook.html).
#'
#' @includeRmd vignettes/codebook.Rmd
#'
#' @format `r pkgsnip::return_label("data")`
#' @aliases codebook
#' @seealso Metadata functions [v_vals()] and [val_lbls()].
#' @export
"data_codebook"
```

### `val_lbls`

```{r}
#' Get the set of possible *value labels* of a referendum data variable
#'
#' Returns a character vector of value labels of a specific [referendums()] column, in the same order as [v_vals()], or of length `0` if `v_name`'s values are
#' not restricted to a predefined set or no value labels are defined in the [codebook][data_codebook].
#'
#' @param v_name A variable name. Must be one of the column names of [`data_codebook`].
#' @param incl_affixes Whether or not to add the corresponding `value_label_prefix` and `value_label_suffix` to the returned labels.
#'
#' @return A character vector. Of length `0` if `v_name`'s values are not restricted to a predefined set or no value labels are defined in the
#'   [codebook][data_codebook].
#' @seealso [`data_codebook`][data_codebook]
#' @family metadata
#' @export
#'
#' @examples
#' c2d::val_lbls("result",
#'               incl_affixes = FALSE)
#' c2d::val_lbls("result")
#' 
#' # Convert the labels to sentence case with trailing dot
#' c2d::val_lbls("result") |> pal::sentenceify()
val_lbls <- function(v_name,
                     incl_affixes = TRUE) {
  v_name <- rlang::arg_match(arg = v_name,
                             values = data_codebook$variable_name)
  metadata <- data_codebook %>% dplyr::filter(variable_name == !!v_name)
  result <- metadata$value_labels %>% purrr::flatten_chr()
  
  if (incl_affixes) {
    if (!is.na(metadata$value_label_prefix)) result <- paste(metadata$value_label_prefix, result)
    if (!is.na(metadata$value_label_suffix)) result <- paste(metadata$value_label_suffix, result)
  }
  
  result
}
```

### `v_vals`

```{r}
#' Get the set of possible *values* of a referendum data variable
#'
#' Returns a vector of the possible predefined values a specific column in [referendums()] can hold. If the variable values aren't restricted to a predefined
#' set, `NULL` is returned.
#'
#' @param v_name A variable name. Must be one of the column names of [`data_codebook`].
#'
#' @return
#' If `v_name`'s values are restricted to a predefined set and
#' - `v_name` is *not* of type list, a vector of the same type as `v_name`.
#' - `v_name` is of type list, a vector of the same type as the elements of `v_name`.
#'
#' Else `NULL`.
#' @seealso [`data_codebook`][data_codebook]
#' @family metadata
#' @export
#'
#' @examples
#' v_vals("result")
#' v_vals("id")
v_vals <- function(v_name) {
  
  v_name <- rlang::arg_match(arg = v_name,
                             values = data_codebook$variable_name)
  data_codebook %>%
    dplyr::filter(variable_name == !!v_name) %$%
    variable_values %>%
    unlist()
}
```

## `count_referendums`

```{r}
#' Count number of referendums
#'
#' Counts the number of referendums in the C2D Database.
#'
#' @inheritParams assemble_query_filter
#'
#' @return A named list with `level` as names and referendum counts as values.
#' @export
#'
#' @examples
#' # the whole database (excl. drafts)
#' c2d::count_referendums()
#' 
#' # only Swiss and Austrian referendums
#' c2d::count_referendums(country_code = c("CH", "AT"))
#' 
#' # only Swiss referendums created between 2020 and 2021
#' c2d::count_referendums(country_code = "CH",
#'                        date_time_created_min = "2020-01-01",
#'                        date_time_created_max = "2021-01-01")
count_referendums <- function(is_draft = FALSE,
                              country_code = NULL,
                              subnational_entity_name = NULL,
                              municipality = NULL,
                              level = NULL,
                              type = NULL,
                              date = NULL,
                              date_time_created_min = NULL,
                              date_time_created_max = NULL,
                              query_filter = NULL) {
  httr::RETRY(verb = "GET",
              url = "https://services.c2d.ch/referendums/stats",
              query = list(filter = assemble_query_filter(country_code = country_code,
                                                          subnational_entity_name = subnational_entity_name,
                                                          municipality = municipality,
                                                          level = level,
                                                          type = type,
                                                          date = date,
                                                          is_draft = is_draft,
                                                          date_time_created_min = date_time_created_min,
                                                          date_time_created_max = date_time_created_max,
                                                          query_filter = query_filter)),
              times = 5L) %>%
    httr::content(as = "parsed") %$%
    votes %>%
    magrittr::set_names(names(.) %>% dplyr::recode("sub_national" = "subnational"))
}
```

## `download_file_attachment`

```{r}
#' Download file attachment
#'
#' Downloads a file attachment from the C2D database. The necessary `s3_object_key`s identifying individual files are found in the `files` list column returned
#' by [referendums()].
#'
#' @param s3_object_key The key uniquely identifying the file in the C2D [Amazon S3 bucket](https://en.wikipedia.org/wiki/Amazon_S3#Design). A character scalar.
#' @param path The path where the downloaded file is written to. If a directory, the original filename returned by the C2D services API will be used.
#'
#' @return A [response object][httr::response], invisibly.
#' @export
#'
#' @examples
#' \dontrun{
#' # get all file object keys...
#' c2d::referendums()$files |>
#'   purrr::map_depth(2L, purrr::pluck, "s3_object_key") |>
#'   purrr::flatten() |>
#'   purrr::flatten_chr() |>
#'   # ...select first three keys...
#'   magrittr::extract(1:3) |>
#'   # ...and download the corresponding files to the current working dir
#'   purrr::walk(c2d::download_file_attachment)}
download_file_attachment <- function(s3_object_key,
                                     path = ".") {
  
  checkmate::assert_string(s3_object_key)
  checkmate::assert_atomic(path)
  
  if (fs::dir_exists(path)) {
    checkmate::assert_directory(path,
                                access = "rw")
    use_original_filename <- TRUE
  } else {
    checkmate::assert_path_for_output(path,
                                      overwrite = TRUE)
    use_original_filename <- FALSE
  }
  
  temp_path <- fs::file_temp()
  
  response <- httr::RETRY(verb = "GET",
                          url = paste0("https://services.c2d.ch/s3_objects/", s3_object_key),
                          httr::write_disk(path = temp_path),
                          times = 5L)
  
  if (use_original_filename) {
    
    final_path <-
      response %>%
      httr::headers() %$%
      `content-disposition` %>%
      stringr::str_extract(pattern = "(?<=filename=\").+?(?=\")") %>%
      fs::path(path, .)
    
  } else {
    final_path <- path
  }
  
  fs::file_move(path = temp_path,
                new_path = final_path)
  
  invisible(response)
}
```

## Manipulate data

TODO

## `search_referendums`

Currently, the C2D API [only supports searching in the English referendum
title](https://github.com/ccmdesign/c2d-app/blob/master/docs/services.md#3-referendum-routes).

```{r}
#' Search in English referendum titles
#'
#' Allows to use the C2D API's primitive search functionality. The search is not case-sensitive and no [fuzzy
#' search](https://en.wikipedia.org/wiki/Approximate_string_matching) is performed (i.e. only exact matches are returned).
#'
#' Note that this function is probably not of much use since it doesn't return any additional information about the matched referendums but only the English
#' titles.
#'
#' @param term The Search term. A character scalar.
#'
#' @return A character vector of English referendum titles matching the search `term`.
#' @export
#'
#' @examples
#' c2d::search_referendums("freedom")
search_referendums <- function(term) {
  
  httr::RETRY(verb = "GET",
              url = "https://services.c2d.ch/referendums",
              query = list(mode = "search",
                           term = checkmate::assert_string(term)),
              times = 5L) %>%
    httr::content(as = "parsed") %$%
    items %>%
    purrr::flatten_chr() 
}
```

## Tags-related

### `hierarchize_tags`

```{r}
#' Hierarchize Tags
#'
#' Reconstructs the hierarchical relations between the three tag variables `tags_tier_1`, `tags_tier_2` and `tags_tier_3`. Can also be used to simply determine
#' the parent tag(s) of any tag.
#'
#' @param x The tags to hierarchize. Either a character vector of tags or a single-row data frame containing at least the columns `tags_tier_1`, `tags_tier_2`
#'   and `tags_tier_3`.
#'
#' @return `r pkgsnip::param_label("data")`
#' @family tags
#' @export
#'
#' @examples
#' c2d::hierarchize_tags("territorial questions")
#'
#' # hierarchize the tags of all Austrian referendums
#' c2d::referendums(country_code = "AT") |>
#'   dplyr::group_by(id) |>
#'   dplyr::group_split() |>
#'   purrr::map(c2d::hierarchize_tags)
hierarchize_tags <- function(x) {
  
  test_char <- checkmate::test_character(x,
                                         any.missing = FALSE)
  
  if (!test_char) {
    
    tag_var_names <- paste0("tags_tier_", 1:3)
    test_df <- checkmate::test_data_frame(x,
                                          min.rows = 1L,
                                          max.rows = 1L)
    has_tag_vars <- all(tag_var_names %in% colnames(x))
    
    if (!test_df || !has_tag_vars) {
      cli::cli_abort(paste0("{.arg x} must be either a character vector of tags or a single-row data frame containing at least the columns ",
                            "{.field tags_tier_1}, {.field tags_tier_2} and {.field tags_tier_3}."))
    }
    
    x %<>%
      .[, tag_var_names] %>%
      purrr::pmap(~ as.character(c(..1, ..2, ..3))) %>%
      purrr::flatten_chr()
  }
  
  checkmate::assert_subset(x,
                           choices = tags_cleaned,
                           empty.ok = TRUE,
                           .var.name = "tags")
  
  tags_tier_1 <- x[x %in% tags_tier_1_cleaned]
  tags_tier_2 <- x[x %in% tags_tier_2_cleaned]
  tags_tier_3 <- x[x %in% tags_tier_3_cleaned]
  inferred_tags_tier_1 <- infer_tags(tags = c(tags_tier_2, tags_tier_3),
                                     tier = 1L)
  inferred_tags_tier_2 <- infer_tags(tags = tags_tier_3,
                                     tier = 2L)
  non_parent_tags_tier_1 <- tags_tier_1 %>% setdiff(inferred_tags_tier_1)
  non_parent_tags_tier_2 <- tags_tier_2 %>% setdiff(inferred_tags_tier_2)
  
  # 0. initialize empty tibble
  result <- tibble::tibble(tag_tier_1 = character(),
                           tag_tier_2 = character(),
                           tag_tier_3 = character())
  
  # 1. add third-tier tags
  result <-
    tags_tier_3 %>%
    purrr::map_dfr(~ tibble::tibble(tag_tier_1 = infer_tags(tags = .x,
                                                            tier = 1L),
                                    tag_tier_2 = infer_tags(tags = .x,
                                                            tier = 2L),
                                    tag_tier_3 = .x)) %>%
    dplyr::bind_rows(result)
  
  # 2. add remaining second-tier tags
  if (length(non_parent_tags_tier_2)) {
    
    result <-
      non_parent_tags_tier_2 %>%
      purrr::map_dfr(~ tibble::tibble(tag_tier_1 = infer_tags(tags = .x,
                                                              tier = 1L),
                                      tag_tier_2 = .x,
                                      tag_tier_3 = NA_character_)) %>%
      dplyr::bind_rows(result)
  }
  
  # 3. add remaining top-tier tags
  if (length(non_parent_tags_tier_1)) {
    
    result %<>% dplyr::bind_rows(tibble::tibble(tag_tier_1 = non_parent_tags_tier_1,
                                                tag_tier_2 = NA_character_,
                                                tag_tier_3 = NA_character_))
  }
  
  # sort result
  result %>% dplyr::arrange(tag_tier_1, tag_tier_2, tag_tier_3)
}
```

### `infer_tags`

```{r}
#' Infer higher-tier tags
#'
#' Determines the top-tier (`tier = 1L`) or second-tier (`tier = 2L`) tags that are the parents or grandparents of `tags` in the
#' [hierarchy](https://rpkg.dev/c2d/articles/codebook.html#tags).
#'
#' @param tags The tags to determine the corresponding (grand)parent tags of. A factor or character vector.
#' @param tier The tier of the inferred tags. Should be >= than the tier of the `tags`. An integer scalar.
#'
#' @return A character vector.
#' @family tags
#' @export
#'
#' @examples
#' c2d::infer_tags(tags = c("EU", "animal protection"),
#'                 tier = 1L)
#' c2d::infer_tags(tags = c("EU", "animal protection"),
#'                 tier = 2L)
#' 
#' # tags of different tiers can mixed in `tags`
#' c2d::infer_tags(tags = c("EU", "environment"),
#'                 tier = 2L)
#' 
#' # but `tags` of a higher tier than `tier` will be ignored
#' c2d::infer_tags(tags = "foreign policy",
#'                 tier = 2L)
infer_tags <- function(tags,
                       tier = 1L) {
  
  if (is.factor(tags)) tags <- as.character(tags)
  
  checkmate::assert_subset(tags,
                           choices = tags_cleaned)
  checkmate::assert_int(tier,
                        lower = 1L,
                        upper = 2L)
  
  # inferred from lower-tier tags
  result <- data_tags_tidy_cleaned[data_tags_tidy_cleaned$tag_tier_2 %in% tags | data_tags_tidy_cleaned$tag_tier_3 %in% tags, ]
  result %<>% .[[paste0("tag_tier_", tier)]]
  
  # plus top-tier tags
  if (tier == 1L) result %<>% c(tags[tags %in% tags_tier_1_cleaned])
  
  unique(result)
}
```

## `validate_referendums`

This function performs various data validation steps to ensure there are no errors in the database.

```{r}
#' Validate referendums
#'
#' @param data The referendum data to validate, as returned by [referendums()].
#' @param check_applicability_constraint Whether or not to check that no applicability constraints as defined in the [codebook][data_codebook] are violated.
#' @param check_id_sudd_prefix Whether or not to check that all [`id_sudd`](https://rpkg.dev/c2d/articles/codebook.html#id-sudd) prefixes are valid.
#'
#' @return `data`, invisibly.
#' @export
validate_referendums <- function(data,
                                 check_applicability_constraint = TRUE,
                                 check_id_sudd_prefix = TRUE) {
  
  checkmate::assert_data_frame(data,
                               min.rows = 1L)
  checkmate::assert_subset(colnames(data),
                           choices = data_codebook$variable_name)
  checkmate::assert_flag(check_applicability_constraint)
  checkmate::assert_flag(check_id_sudd_prefix)
  
  # check applicability constraints
  if (check_applicability_constraint) {
    
    v_names_violated <-
      data_codebook %>%
      dplyr::filter(variable_name %in% colnames(data)
                    & !is.na(applicability_constraint)) %$%
      purrr::map2_lgl(.x = magrittr::set_names(x = variable_name,
                                               value = variable_name),
                      .y = applicability_constraint,
                      .f = ~ {
                        
                        data %>%
                          dplyr::filter(!eval(parse(text = .y))) %$%
                          eval(as.symbol(.x)) %>%
                          {is.na(.) | purrr::map_lgl(., is.null)} %>%
                          all()
                      }) %>%
    magrittr::extract(. %in% FALSE) %>%
    names()
    
    n_v_names_violated <- length(v_names_violated)
    
    if (n_v_names_violated) {
      
      cli::cli_alert_warning("Applicability constraints are violated for {n_v_names_violated} variable{?s}:")
      
      paste0("{.var ", v_names_violated, "}") %>%
        magrittr::set_names(rep("x",
                                times = length(.))) %>%
        cli::cli_bullets()
      
      first_v_name_violated <- v_names_violated[1L]
      
      cli::cli_text("To get the applicability constraint of e.g. {.var {first_v_name_violated}}, run:")
      cat("\n")
      cli::cli_code(c("data_codebook %>%",
                      glue::glue("  dplyr::filter(variable_name == \"{first_v_name_violated}\") %$%"),
                      "  applicability_constraint"))
      cat("\n")
      cli::cli_text("To inspect the entries in violation of the above applicability constraint, run:")
      cat("\n")
      cli::cli_code(c("data_codebook %>%",
                      glue::glue("  dplyr::filter(variable_name == \"{first_v_name_violated}\") %$%"),
                      "  dplyr::filter(.data = data,",
                      "                !eval(parse(text = applicability_constraint))) %$%",
                      glue::glue("  {first_v_name_violated}")))
    }
  }
  
  # check `id_sudd` prefix if requested
  if (check_id_sudd_prefix) {
    
    if (!all(c("country_code", "id_sudd") %in% colnames(data))) {
      cli::cli_abort("Columns {.var country_code} and {.var id_sudd} must be present in {.arg data}.")
    }
    
    # define allowed exceptions
    allowed_exceptions <- tibble::tribble(
      ~country_code, ~id_sudd_prefix,
      # Curaao
      "CW", "an",
      # Szeklerland, cf. https://sudd.ch/event.php?id=hu042008
      "RO", "hu"
    )
    
    # assemble target country codes
    country_codes <- data$country_code %>% as.list()
    
    for (country_code in allowed_exceptions$country_code) {
      
      additional_country_codes <-
        allowed_exceptions %>%
        dplyr::filter(country_code == !!country_code) %$%
        id_sudd_prefix %>%
        stringr::str_to_upper()
      i_country_codes <- country_codes %>% purrr::map_lgl(~ country_code == .x) %>% which()
      
      for (i in i_country_codes) {
        country_codes[[i]] <- unique(c(country_codes[[i]], additional_country_codes))
      }
    }
    
    # add dummy indicating if target country codes match
    # TODO: instead of modifying input data, print cli msg with all relevant info!
    data$matches_id_sudd_prefix <-
      data$id_sudd %>%
      stringr::str_extract(pattern = "^..") %>%
      stringr::str_to_upper() %>%
      purrr::map2_lgl(.y = country_codes,
                      .f = ~ .x %in% .y)
    
    data$matches_id_sudd_prefix[is.na(data$id_sudd)] <- NA
  }
  
  invisible(data)
}
```
