---
title: "C2D Codebook"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{C2D Codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Referendum-level variables

For the following variables, the [unit of observation](https://en.wikipedia.org/wiki/Unit_of_observation) is the individual referendum.

```{r, referendum-level-vars, echo = FALSE, results = "asis", message = FALSE}
# setup
library(rlang)
library(magrittr)

# define functions
## flatten list column values
flatten_list_col <- function(x) {
  
  x %>%
    unlist() %>%
    purrr::when(is.character(.) ~ pal::wrap_chr(.),
                ~ .)
}

## convert a string to sentence-case with trailing dot
sentencify <- function(x) {
  
  x %>%
    stringr::str_replace(pattern = "^.",
                         replacement = toupper) %>%
    purrr::map_chr(~ {if (stringr::str_sub(string = .x, start = -1L) == ".") .x else paste0(.x, ".")})
}

## prettify value scale
prettify_value_scale <- function(x) {
  
  x %>%
    stringr::str_replace(pattern = "_(.*)",
                         replacement = " (\\1)") %>%
    dplyr::recode("binary" = "nominal (binary)")
}

## assemble variable sections in Markdown
cat_v_sections <- function(data) {
  
  data %>% purrr::pwalk(function(variable_name,
                                 variable_label,
                                 is_multi_valued,
                                 applicability_constraint,
                                 variable_values,
                                 value_label_prefix,
                                 value_labels,
                                 value_label_suffix,
                                 value_default,
                                 value_scale) {
    
    variable_values %<>% flatten_list_col()
    value_default %<>% flatten_list_col()
    value_scale %<>% prettify_value_scale()
    value_labels %<>%
      unlist() %>%
      purrr::when(is.na(value_label_prefix) ~ .,
                  ~ paste(value_label_prefix, .)) %>%
      purrr::when(is.na(value_label_suffix) ~ .,
                  ~ paste(., value_label_suffix)) %>%
      sentencify()
    
    pal::cat_lines(glue::glue("### `{variable_name}`"),
                   "",
                   variable_label %>% sentencify(),
                   "",
                   if (length(variable_values)) {
                     c(glue::glue("#### Possible values {{#{variable_name}-values}}"),
                       "",
                       if (length(value_labels)) {
                         c("| Value | Meaning |",
                           "| :---- | :------ |",
                           variable_values %>% purrr::map2_chr(.y = value_labels,
                                                               .f = ~ glue::glue("| `{.x}` | {.y} |")))
                       } else {
                         c("| Value |",
                           "| :---- |",
                           variable_values %>% purrr::map_chr(~ glue::glue("| `{.x}` |")))
                       },
                       # if (!is.null(value_default)) c("", glue::glue("The default value is: `{value_default}`")),
                       "")
                   },
                   if (is_multi_valued) {
                     c("_This is a multi-value variable[^multi-val-var] and thus of type `list`._",
                       "")
                   },
                   if (!is.na(applicability_constraint)) {
                     c(glue::glue("#### Applicability constraint {{#{variable_name}-constraint}}"),
                       "",
                       "This variable is only applicable under the following condition:",
                       "",
                       "```r",
                       applicability_constraint,
                       "```",
                       "")
                   },
                   if (value_scale != "undefined") {
                     c(glue::glue("#### Level of measurement {{#{variable_name}-scale}}"),
                       "",
                       glue::glue("This variable's [scale of measure](https://en.wikipedia.org/wiki/Level_of_measurement) is the **{value_scale}** level."),
                       "")
                   })
  })
}

c2d:::data_codebook %>%
  dplyr::filter(stringr::str_detect(string = variable_name,
                                    pattern = "^inst_",
                                    negate = TRUE)) %>%
  cat_v_sections()
```

## Institution-level variables

For the following variables, the [unit of observation](https://en.wikipedia.org/wiki/Unit_of_observation) is the referendum's *institutional instrument*, a
combination of the referendum's `type` and the jurisdiction it took place.

The set of variables that constitute the jurisdiction is dependent on the referendum's `level`:

| `level`       | Jurisdiction is given by                                    |
|:--------------|:------------------------------------------------------------|
| `national`    | `country_code`                                              |
| `subnational` | `country_code` + `subnational_entity_name`                  |
| `local`       | `country_code` + `subnational_entity_name` + `municipality` |

```{r, institution-level-vars, echo = FALSE, results = "asis"}
c2d:::data_codebook %>%
  dplyr::filter(stringr::str_detect(string = variable_name,
                                    pattern = "^inst_")) %>%
  cat_v_sections()
```

```{r, footnotes, echo = FALSE, results = "asis"}
cat("
[^multi-val-var]: A multi-value variable allows for more than one value per observation and thus breaks with the [tidy-data convention](https://tidyr.tidyverse.org/articles/tidy-data.html). Use [`tidyr::unnest_longer()`](https://tidyr.tidyverse.org/reference/hoist.html) to convert data containing such a variable to long format. To unnest `tags` to `tag` for example, use:

    \`\`\`r
    c2d::referendums() %>% tidyr::unnest_longer(col = tags, values_to = \"tag\")
    \`\`\`
")
```
