---
title: "Introduction to rdb"
output:
  rmarkdown::html_vignette:
    css: custom.css
vignette: >
  %\VignetteIndexEntry{Introduction to rdb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

# set RNG seeds
set.seed(42L)
htmlwidgets::setWidgetIdSeed(seed = 42L)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)

# disable cache expiry
options(rdb.global_max_cache_age = Inf)

# set pkg metadata
pkg_metadata <- pal::desc_list()
```

# Installation

rdb is not on CRAN yet[^1], but you can install the package directly from the development source code repository's [`master`
branch](https://gitlab.com/zdaarau/rpkgs/rdb/-/tree/master), which we try to keep in a working state at all times.

```{r}
#| label: install-snippet-dev
#| child: !expr pkgsnip::snip_path("installation-notice_dev-version_gitlab.Rmd")
```

[^1]: We will probably release the package on CRAN once the current database and backend overhaul is complete and the rdb R package is properly adapted to use
    the new infrastructure, as well as sufficiently tested. This will most likely not happen before the end of 2024.

# Usage

## Get data

You can download RDB referendum data via the two functions `rdb::rfrnd()` and `rdb::rfrnds()`. The former one fetches the data of a single referendum only, of
which you must already know its uniqe RDB [`id`](https://rdb.rpkg.dev/articles/codebook.html#id). The latter function allows to retrieve data for an arbitrary
number of referendums, depending on the conditions you specify via the [function's various arguments](https://rdb.rpkg.dev/reference/rfrnds.html#arguments).

To simply retrieve *all* referendums in the database (excluding draft entries), run

``` r
rdb::rfrnds()
```

which should output a [tibble](https://tibble.tidyverse.org/) like this one:

```{r}
#| echo: false

rdb::rfrnds()
```

The RDB referendum data's individual variables (columns) are documented in the [codebook](https://rdb.rpkg.dev/articles/codebook.html). It is also available as
a dataset via `rdb::data_codebook`.

Results of `rdb::rfrnds()` and some other functions in this package are by default cached on disk using [pkgpins](https://pkgpins.rpkg.dev/)[^2]. You can define
the maximum age of cached results you're willing to tolerate via the argument `max_cache_age` (defaults to a week). It accepts anything that can be successfully
converted to a [lubridate duration](https://lubridate.tidyverse.org/reference/as.duration.html) -- e.g. a string like `"3 hours"`, `"2 days"` or `"1 week"`, or
a number which will simply be interpreted as *number of seconds*.

To only re-download RDB data once every 4 hours and 48 minutes for example, use

``` r
rdb::rfrnds(max_cache_age = "4 hours 48 minutes")
```

Although we usually advise against it, you can also completely opt out of caching by specifying `use_cache = FALSE`. However, please make sure to not run such
code in excess, as it creates additional (and most likely unnecessary) load on our servers.

[^2]: Caching happens per unique combination of *evaluated* function arguments (excluding any of the arguments `incl_archive`, `use_cache`, `max_cache_age` and
    `quiet`). This means that when `rdb::rfrnds()` is invoked two times in a row with the exact same set of arguments, say `country_code = "CH"` and
    `level = c("national", "subnational")`, the second invocation will read the result from disk and thus is considerably faster (as long as `use_cache` is set
    to `TRUE` and `max_cache_age` is set to a timespan greater than the delay between invocations).

    Since every argument combination is cached separately, altering the arguments will inevitably result in a new cache entry, regardless whether the actual
    data would technically already be present in an existing cache entry. E.g. `rdb::rfrnds(country_code = "CH", level = "national")` returns a perfect subset
    of the just mentioned invocation, nevertheless the data is freshly fetched the first time it is run.

## Augment data

rdb includes various functions to augment the referendum data by additional information which wouldn't make sense to be stored in the RDB itself.

For example, you can add the period (week, month, quarter, year, decade or century) in which a referendum took place using `rdb::add_period()`. By default, the
recurring numeric week number of the year is added (i.e. `period = "week"`):

```{r}
#| echo: true

rdb::rfrnds() |>
  rdb::add_period() |>
  dplyr::select(id, date, week)
```

Another frequently required augmentation is `rdb::add_country_code_long()` which adds an additional column `country_code_long` containing the [ISO 3166-1
alpha-3](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3) code. These three-letter codes are often required to join RDB referendum data with data from other
sources.

See the package reference for [all available data augmentation functions](https://rdb.rpkg.dev/reference/index.html#augmentation).

## Transform data

For certain analyses, it might come in handy to transform the referendum data to a different shape beforehand. For a few such transformations, rdb provides
ready-made functions.

`rdb::as_ballot_dates()` for example transforms the default referendum-level observations to ones on the level of ballot date and jurisdiction:

```{r}
#| echo: true

rdb::rfrnds() |> nrow()

rdb::rfrnds() |> rdb::as_ballot_dates() |> nrow()
```

Noteworthy is also `rdb::unnest_var()` which provides a convenient and standardized way to unnest a multi-value variable of type list like the [`topics_tier_*`
variables](https://rdb.rpkg.dev/articles/codebook.html#topics) to long format.

```{r}
#| echo: true

rdb::rfrnds() |> nrow()

rdb::rfrnds() |> rdb::unnest_var(topics_tier_1) |> nrow()
```

See the package reference for [all available data transformation functions](https://rdb.rpkg.dev/reference/index.html#transformation).

## Tabulate and visualize data

rdb also includes some ready-made convenience functions to create tables and (interactive) plots.

If you'd like a tabular overview of the top-ten countries by number of ballot dates per political level for example, you could simply run

``` r
rdb::rfrnds() |>
  rdb::as_ballot_dates() |>
  rdb::tbl_n_rfrnds(by = c(country_name, level),
                    n_rows = 10L,
                    order = "descending")
```

and you'd get the following nicely formatted [gt](https://gt.rstudio.com/) table:

```{r}
#| echo: false

rdb::rfrnds() |>
  rdb::as_ballot_dates() |>
  rdb::tbl_n_rfrnds(by = c(country_name, level),
                    n_rows = 10L,
                    order = "descending")
```

A table of the number of referendums in the [UN subregion](https://rdb.rpkg.dev/reference/add_world_regions.html) *Polynesia* since 2010 per a certain period,
say years, can be generated via

``` r
rdb::rfrnds() |>
  rdb::add_world_regions() |>
  dplyr::filter(un_subregion == "Polynesia" & date > "2009-12-31") |>
  rdb::tbl_n_rfrnds_per_period(period = "year")
```

```{r}
#| echo: false

rdb::rfrnds() |>
  rdb::add_world_regions() |>
  dplyr::filter(un_subregion == "Polynesia" & date > "2009-12-31") |>
  rdb::tbl_n_rfrnds_per_period(period = "year")
```

Or a stacked area chart visualizing the worldwide share of referendums per year since 1950, grouped by [political
level](https://rdb.rpkg.dev/articles/codebook.html#level):

``` r
rdb::rfrnds() |>
  dplyr::filter(date >= "1950-01-01") |>
  rdb::tbl_n_rfrnds_per_period(period = "year",
                               by = "level")
```

```{r}
#| echo: false

rdb::rfrnds() |>
  dplyr::filter(date >= "1950-01-01") |>
  rdb::plot_rfrnd_share_per_period(period = "year",
                                   by = "level") |>
  plotlee::simplify_trace_ids()
```

Or, as a final example, the overall (hierarchical) segmentation of the [political topics](https://rdb.rpkg.dev/articles/codebook.html#topics) all the
referendums in the RDB were about:

``` r
rdb::rfrnds() |> rdb::plot_topic_segmentation(method = "per_topic_lineage")
```

```{r}
#| echo: false

rdb::rfrnds() |>
  rdb::plot_topic_segmentation(method = "per_topic_lineage") |>
  plotlee::simplify_trace_ids()
```

Again, see the package reference for [all available data visualization and tabulation functions](https://rdb.rpkg.dev/reference/index.html#visualization). More
will likely be added in the future.
